### 2021-09-13 paper/header.tex
Basic paper structure (juliacon)

% **************GENERATED FILE, DO NOT EDIT**************

\title{CompositionalNetworks.jl: a scaling glass-box neural network for combinatorial functions learning}

\author[1]{Jean-François \textsc{Baffier}}
\author[2]{Khalil \textsc{Chrit}}
\author[34]{Florian \textsc{Richoux}}
\author[2]{Pedro \textsc{Patinho}}
\author[2]{Salvador \textsc{Abreu}}
\affil[1]{IIJ Innovation Institute, Japan}
\affil[2]{NOVA-LINCS, University of Évora, Portugal}
\affil[3]{AIST, Japan}
\affil[4]{JFLI, CNRS, Japan}

\keywords{Julia, Optimization, Constraint Programming, Neural Network, Scaling Learning}

\hypersetup{
pdftitle = {CompositionalNetworks.jl: a scaling glass-box neural network for combinatorial functions learning},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {Jean-François \textsc{Baffier}, Khalil \textsc{Chrit}, Florian \textsc{Richoux}, Pedro \textsc{Patinho}, Salvador \textsc{Abreu}},
pdfkeywords = {Julia, Optimization, Constraint Programming, Neural Network, Scaling Learning},
}

### 2021-09-13 paper/paper.tex
Basic paper structure (juliacon)

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}

\begin{document}

\input{header}

\maketitle

\begin{abstract}

Interpretable Compositional Networks (ICN), a variant of neural networks, that allows the user to get interpretable results, unlike regular artificial neural networks. An ICN is a glass-box producing functions composition that scale with the size of the input, allowing a learning phase on relatively small spaces. This presentation covers the different Julia packages and paradigms involved, a set of use-case, current limitations, future developments, and hopefully possible collaborations.

\end{abstract}

\begin{itemize}
    \item introduction to
    \begin{itemize}
        \item CP
        \item Metaheuristics
        \item CBLS and EFN
    \end{itemize}
    \item Pros of doing things in Julia
    \item paper organization/content
\end{itemize}

\section{Introduction}

The \LaTeX{} document preparation system is a special version of the
\TeX; typesetting program where-in a collection of \TeX; commands
are added to \LaTeX{} to simplify typesetting. Importantly, it allows
the author to concentrate on the logical structure of the document
rather than its visual layout.\vskip 6pt
Moreover, \LaTeX{} provides a consistent and comprehensive document
preparation interface. There are simple-to-use commands for
generating a table of contents, lists of figures and/or tables, and indexes.
\LaTeX{} can automatically number list entries, equations, figures,
tables, and footnotes, as well as articles, sections, and subsections.
Using this numbering system, bibliographic citations, page
references, and cross references to any other numbered entity (e.g.
article, section, equation, figure, list entry, etc.) become quite simple
and straightforward. The use of \LaTeX{} document classes allows
a simple change of class to transform the appearance of your document.\vskip 6pt
\LaTeX{} is a powerful tool for managing long and complex documents.
In particular, partial processing enables long documents to
be produced article by article without losing sequential information.
The use of document classes allows a simple change of style
(or style option) to transform the appearance of your document.

\section{The JuliaCon Article Class}
\label{sec:documentclass}
%
The juliacon class file preserves the standard LATEX{} interface such
that any document that can be produced using the standard LATEX{}
article class can also be produced with the class file.\vskip 6pt
It is likely that the make up will change after file submission. For
this reason, we ask you to ignore details such as slightly long lines,
page stretching, or figures falling out of synchronization, as these
details can be dealt with at a later stage.\vskip 6pt
Use should be made of symbolic references (\verb|\ref| ) in order to
protect against late changes of order, etc.

\section{USING THE JuliaCon Article CLASS FILE}

If the file \verb|juliacon.cls|  is not already in the appropriate system directory
for \LaTeX{} files, either arrange for it to be put there or copy
it to your working directory. The \verb|juliacon|  document class is implemented
as a complete class, not a document style option. In order to
use the \verb|juliacon  document class, replace \verb|article|  by \verb|juliacon|  in the
\verb|\documentclass|  command at the beginning of your document:
\vskip 6pt
\begin{centering}
    \verb|\documentclass{article}|  \end{centering}
\vskip 6pt
replace by
\vskip 6pt
 \verb|\documentclass{juliacon}|  \vskip 6pt
In general, the following standard document \verb|style|  options should
{ \itshape not} be used with the {\footnotesize \itshape article} class file:
\begin{enumerate}
\item[(1)] \verb|10pt|,  \verb|11pt|,  \verb|12pt|   ? unavailable;
\item[(2)] \verb|twoside|  (no associated style file) ? \verb|twoside|  is the default;
\item[(3)] \verb|fleqn|, \verb|leqno|, \verb|titlepage| ? should not be used;
\end{enumerate}

\section{Additional Document Style Options}
\label{sec:additional_doc}
%
The following additional style option is available with the \verb|juliacon|  class file:
\vskip 6pt
Please place any additional command definitions at the very start of
the \LaTeX{} file, before the \verb|\begin{document}|. For example, user-defined
\verb|\def|  and \verb|\newcommand|   commands that define macros for
technical expressions should be placed here. Other author-defined
macros should be kept to a minimum.
\vskip 6pt
Commands that differ from the standard \LaTeX{} interface, or that
are provided in addition to the standard interface, are explained in
this guide. This guide is not a substitute for the \LaTeX{} manual itself.
Authors planning to submit their papers in \LaTeX{} are advised to use
\verb|juliacon.cls|  as early as possible in the creation of their files.

%
%
%
%
\begin{table*}[t]
\tabcolsep22pt
\tbl{If necessary, the tables can be extended both columns.}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Test 1 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test 2 & Testing Data I & & 105\\\hline
Test 3 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\label{tab:symbols}
\begin{tabnote}
This is an example of table footnote.
\end{tabnote}
\end{table*}
% \begin{figure*}[t]
% \centerline{\includegraphics[width=11cm]{juliagraphs.png}}
% \caption{If necessary, the images can be extended both columns.}
%   \label{fig:sample_image}
% \end{figure*}

\section{Additional features}
\label{sec:additional_faci}
In addition to all the standard \LaTeX{} design elements, the \verb|juliacon|  class file includes the following features:
In general, once you have used the additional \verb|juliacon.cls| facilities
in your document, do not process it with a standard \LaTeX{} class
file.

\subsection{Titles, Author's Name, and Affiliation}
\label{subsub:title_auth}
The title of the article, author's name, and affiliation are used at the
beginning of the article (for the main title). These can be produced
using the following code:

\begin{verbatim}
\title{ This is an example of article title} }
\author{
   \large 1st Author \\[-3pt]
   \normalsize 1st author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	1st author's email address \\[-3pt]
  \and
   \large 2nd Author \\[-3pt]
   \normalsize 2nd author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	2nd author's email address \\[-3pt]
\and
   \large 3rd Author \\[-3pt]
   \normalsize 3rd author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	3rd author's email address \\[-3pt]
}
\maketitle
\end{verbatim}

\subsection{Writing Julia code}

A special environment is already defined for Julia code,
built on top of \textit{listings} and \textit{jlcode}.

\begin{verbatim}
\begin{lstlisting}[language = Julia]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}
\end{verbatim}
\begin{lstlisting}[language = Julia]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}

\subsection{Abstracts, Key words, term etc...}
\label{subsub:abs_key_etc}

At the beginning of your article, the title should be generated
in the usual way using the \verb|\maketitle|  command. For genaral tem and keywords use
\verb|\terms|,
\verb|\keywords|  commands respectively. The abstract should be enclosed
within an abstract environment, All these environment
can be produced using the following code:
\begin{verbatim}
\terms{Experimentation, Human Factors}

\keywords{Face animation, image-based modelling...}

\begin{abstract}
In this paper, we propose a new method for the
systematic determination of the model's base of
time varying delay system. This method based on
the construction of the classification data related
to the considered system. The number, the orders,
the time delay and the parameters of the local
models are generated automatically without any
knowledge about the full operating range of the
process. The parametric identification of the local
models is realized by a new recursive algorithm for
on line identification of systems with unknown time
delay. The proposed algorithm allows simultaneous
estimation of time delay and parameters of
discrete-time systems. The effectiveness of
the new method has been illustrated through
simulation.
\end{abstract}

\end{verbatim}

\section{Some guidelines}
\label{sec:some_guide}
The following notes may help you achieve the best effects with the
\verb|juliacon|  class file.

\subsection{Sections}
\label{subsub:sections}
\LaTeXe{}  provides four levels of section headings and they are all
defined in the \verb|juliacon|  class file:
\begin{itemize}
\item \verb|\section|
\item \verb|\subsection|
\item \verb|\subsubsection|
\item \verb|\paragraph|
\end{itemize}
Section headings are automatically converted to allcaps style.
\subsection{Lists}
\label{sec:lists}
%
The \verb|juliacon|   class file provides unnumbered lists using the
\verb|unnumlist|  environment for example,

\begin{unnumlist}
\item First unnumbered item which has no label and is indented from the
left margin.
\item Second unnumbered item.
\item Third unnumbered item.
\end{unnumlist}
The unnumbered list which has no label and is indented from the
left margin. was produced by:
\begin{verbatim}
\begin{unnumlist}
\item First unnumbered item...
\item Second unnumbered item...
\item Third unnumbered item...
\end{unnumlist}
\end{verbatim}

The \verb|juliacon|  class file also provides hyphen list using the
\verb|itemize|  environment for example,
\begin{itemize}
\item First unnumbered bulleted item which has no label and is indented
from the left margin.
\item Second unnumbered bulleted item.
\item Third unnumbered bulleted item which has no label and is indented
from the left margin.
\end{itemize}
was produced by:
\begin{verbatim}
\begin{itemize}
\item First item...
\item Second item...
\item Third item...
\end{itemize}
\end{verbatim}

Numbered list is also provided in acmtog class file using the
enumerate environment for example,
\begin{enumerate}
\item The attenuated and diluted stellar radiation.
\item Scattered radiation, and
\item Reradiation from other grains.
\end{enumerate}

was produced by:
\begin{verbatim}
\begin{enumerate}
\item The attenuated...
\item Scattered radiation, and...
\item Reradiation from other grains...
\end{enumerate}
\end{verbatim}
\subsection{Illustrations (or figures)}
\label{subsub:sec_Illus}
The \verb|juliacon|  class file will cope with most of the positioning of
your illustrations and you should not normally use the optional positional
qualifiers on the \verb|figure|  environment that would override
these decisions.
\vskip 6pt

%
\begin{figure}[t]
\centerline{\includegraphics[width=4cm]{juliagraphs.png}}
\caption{This is example of the image in a column.}
	\label{fig:sample_figure}
\end{figure}

The figure \ref{fig:sample_figure} is taken from the JuliaGraphs
organization \footnote{https://github.com/JuliaGraphs}.

Figure captions should be \emph{below} the figure itself, therefore the
\verb|\caption|  command should appear after the figure or space left for
an illustration. For example, Figure 1 is produced using the following
commands:

\begin{verbatim}
\begin{figure}
\centerline{\includegraphics[width=20pc]{Graphics.eps}}
\caption{An example of the testing process for a
binary tree. The globa null hypothesis is tested
first at level $\alpha$ (a), and the level of
individual variables is reached last (d). Note
that individual hypotheses can be tested at
level $\alpha/4$ and not $\alpha/8$ as one might
expect at first.}
\label{sample-figure_2}
\end{figure}
\end{verbatim}
Figures can be resized using first and second argument of
\verb|\includegraphics|   command. First argument is used for modifying
figure height and the second argument is used for modifying
figure width respectively.
\vskip 6pt
Cross-referencing of figures, tables, and numbered, displayed
equations using the \verb|\label|  and \verb|\ref|   commands is encouraged.
For example, in referencing Figure 1 above, we used
\verb|Figure~\ref{sample-figure}|

 \subsection{Tables}
\label{subsub:sec_Tab}
The \verb|juliacon|   class file will cope with most of the positioning of
your tables and you should not normally use the optional positional qualifiers on the table environment which would override these
decisions. Table captions should be at the top.
\begin{verbatim}
\begin{table}
\tbl{Tuning Set and Testing Set}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Train70 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test70 & Testing Data I & & 105\\\hline
Test30 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\end{table}
\end{verbatim}

\begin{table}
\tbl{Tuning Set and Testing Set}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Test 1 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test 2 & Testing Data I & & 105\\\hline
Test 3 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\end{table}
\subsection{Landscaping Pages}
\label{subsub:landscaping_pages}
If a table is too wide to fit the standard measure, it may be turned,
with its caption, to 90 degrees. Landscape tables cannot be produced
directly using the \verb|juliacon|   class file because \TeX{} itself cannot
turn the page, and not all device drivers provide such a facility.
The following procedure can be used to produce such pages.
\vskip 6pt
Use the package \verb|rotating|   in your document and change the coding
from
\begin{verbatim}
\begin{table}...\end{table}
to
\begin{sidewaystable}...\end{sidewaystable}
and for figures
\begin{figure}...\end{figure}
to
\begin{sidewaysfigure}...\end{sidewaysfigure}
\end{verbatim}

environments in your document to turn your table on the appropriate
page of your document. For instance, the following code prints
a page with the running head, a message half way down and the
table number towards the bottom.
\begin{verbatim}
\begin{sidewaystable}
\tbl{Landscape table caption to go here.}{...}
\label{landtab}
\end{sidewaystable}
\end{verbatim}

\subsection{Double Column Figure and Tables}
\label{subsub:double_fig_tab}
For generating the output of figures and tables in double column
we can use the following coding:

\begin{enumerate}
\item For Figures:
\begin{verbatim}
\begin{figure*}...\end{figure*}
\end{verbatim}
\item For landscape figures:
\begin{verbatim}
\begin{sidewaysfigure*}...\end{sidewaysfigure*}
\end{verbatim}
\item For Tables:
\begin{verbatim}
\begin{table*}...\end{table*}
\end{verbatim}
\item For landscape tables:
\begin{verbatim}
\begin{sidewaystable*}...\end{sidewaystable*}
\end{verbatim}
\end{enumerate}

\subsection{Typesetting Mathematics}
\label{subsub:type_math}
The \verb|juliacon| class file will set displayed mathematics with center to
the column width, provided that you use the \LaTeXe{} standard of
open and closed square brackets as delimiters.
The equation
\[
\sum_{i=1}^p \lambda_i = (S)
\]

was typeset using the acmtog class file with the commands

\begin{verbatim}
\[
\sum_{i=1}^p \lambda_i = (S)
\]
\end{verbatim}

For display equations, cross-referencing is encouraged. For example,
\begin{verbatim}
\begin{equation}
(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
\label{eq:samplevar}
\end{equation}
Equation~(\ref{eq:samplevar}) gives the formula for
sample variance.
\end{verbatim}
The following output is generated with the above coding:
\begin{equation}
(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
\label{eq:samplevar}
\end{equation}
Equation~(\ref{eq:samplevar}) gives the formula for
sample variance.

\subsection{Enunciations}
\label{subsub:enunciation}
The \verb|juliacon|   class file generates the enunciations with the help of
the following commands:
\begin{verbatim}
\begin{theorem}...\end{theorem}
\begin{strategy}...\end{strategy}
\begin{property}...\end{property}
\begin{proposition}...\end{proposition}
\begin{lemma}...\end{lemma}
\begin{example}...\end{example}
\begin{proof}...\end{proof}
\begin{definition}...\end{definition}
\begin{algorithm}...\end{algorithm}
\begin{remark}...\end{remark}
\end{verbatim}
The above-mentioned coding can also include optional arguments
such as
\begin{verbatim}
\begin{theorem}[...]. Example for theorem:
\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven ... created equal.
\end{theorem}
\end{verbatim}

\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven years ago our fathers brought forth,
upon this continent, a new nation, conceived in Liberty,
 and dedicated to the proposition that all men are
created equal.
\end{theorem}

\subsection{Extract}
\label{subsub:extract}
Extract environment should be coded within
\begin{verbatim}
\begin{extract}..\end{extract}
\end{verbatim}

\subsection{Balancing column at last page}
\label{subsub:Balance}
For balancing the both column length at last page use :
\begin{verbatim}
\vadjust{\vfill\pagebreak}
\end{verbatim}

%\vadjust{\vfill\pagebreak}

at appropriate place in your \TeX{} file or in bibliography file.

\section{Handling references}
\label{subsub:references}
References are most easily (and correctly) generated using the
BIBTEX, which is easily invoked via
\begin{verbatim}
\bibliographystyle{juliacon}
\bibliography{ref}
\end{verbatim}
When submitting the document source (.tex) file to external
parties, the ref.bib file should be sent with it.
\cite{bezanson2017julia}

\input{bib.tex}

\end{document}

% Inspired by the International Journal of Computer Applications template

### 2021-09-13 xp/scripts/icn.jl
Write output for icn_benchmark in json files

@@ -18,13 +18,7 @@ using BenchmarkTools
 using CompositionalNetworks
 using ConstraintDomains
 using Constraints
-
-# NOTE - Is write_benchmarks useful?
-# function write_benchmarks(path, data)
-#     file = open(path, "a")
-#     write(file, data)
-#     return close(file)
-# end
+using JSON
 
 function icn_benchmark_unit(params)
     @info "Running a benchmark unit with" params
@@ -54,60 +48,104 @@ function icn_benchmark_unit(params)
             end
         end
     end
-
-    # Generate an appropriate parameter for the concept if relevant
-    param = if isnothing(params[:concept][2])
-        nothing
-    elseif params[:concept][2] == 1
-        rand(1:params[:domains_size])
+    json_name = savename(
+        (
+            con=string(params[:concept][1]),
+            par=params[:concept][2],
+            csl=params[:complete_search_limit],
+            dom=params[:domains_size],
+            gen=params[:generations],
+            iter=params[:icn_iterations],
+            lang=string(params[:language]),
+            lst=params[:loss_sampling_threshold],
+            ls=string(params[:loss_sampler]),
+            metric=string(params[:metric]),
+            psl=params[:partial_search_limit],
+            pop=params[:population],
+            sampling=params[:sampling],
+            search=string(params[:search]),
+        ),
+        "json",
+    )
+    save_results = joinpath(datadir("compositions"), json_name)
+    if isfile(save_results)
+        @warn "The result file already exist" save_results
     else
-        rand(1:params[:domains_size], params[:concept][2])
+
+        # Generate an appropriate parameter for the concept if relevant
+        param = if isnothing(params[:concept][2])
+            nothing
+        elseif params[:concept][2] == 1
+            rand(1:params[:domains_size])
+        else
+            rand(1:params[:domains_size], params[:concept][2])
+        end
+
+        # assign parameters
+        constraint_concept = concept(BENCHED_CONSTRAINTS[params[:concept][1]])
+        metric = params[:metric]
+        domain_size = params[:domains_size]
+        domains = fill(domain(1:domain_size), domain_size)
+        func_name = "icn" * string(constraint_concept)[8:end] * "_" * string(metric)
+        func_path = datadir("compositions", func_name * ".jl")
+
+        # Time the data retrieval/generation
+        t = @timed search_space(
+            domain_size,
+            constraint_concept,
+            param;
+            search=params[:search],
[…]

### 2021-09-13 xp/scripts/icn.jl
Improve icn_benchmark output

@@ -118,7 +118,13 @@ function icn_benchmark_unit(params)
         # Global results
         push!(results, :data => has_data ? :loaded : :explored)
         push!(results, :data_time => t.time)
+        push!(results, :data_bytes => t.bytes)
+        push!(results, :data_gctime => t.gctime)
+        push!(results, :data_gcstats => t.gcstats)
         push!(results, :icn_time => bench.time)
+        push!(results, :icn_bytes => bench.bytes)
+        push!(results, :icn_gctime => bench.gctime)
+        push!(results, :icn_gcstats => bench.gcstats)
         push!(results, :total_time => t.time + bench.time)
         push!(results, :nthreads => Threads.nthreads())
 
@@ -130,7 +136,12 @@ function icn_benchmark_unit(params)
 
             # Code composition
             for lang in (params[:language], :maths)
-                push!(local_results, lang => CompositionalNetworks.code(compo, lang))
+                push!(
+                    local_results,
+                    lang => CompositionalNetworks.code(
+                        compo, lang; name=string(params[:concept][1])
+                    ),
+                )
             end
             push!(local_results, :symbols => CompositionalNetworks.symbols(compo))
 

### 2021-09-13 src/CompositionalNetworks.jl
Added memoization to icn script

@@ -5,9 +5,11 @@ using ConstraintDomains
 using Dictionaries
 using Evolutionary
 using JuliaFormatter
+using Memoization
 using OrderedCollections
 using Random
 using ThreadPools
+using ThreadSafeDicts
 using Unrolled
 
 export Composition

### 2021-09-13 src/genetic.jl

@@ -38,11 +38,13 @@ function _optimize!(
     pop_size,
     iterations;
     samples=nothing,
+    memoize=true,
 )
+    _metric = memoize ? (@memoize Dict memoize_metric(x, X) = metric(x, X)) : metric
+    _bias = memoize ? (@memoize Dict memoize_bias(x) = weigths_bias(x)) : weigths_bias
     fitness =
         w ->
-            loss(solutions, non_sltns, icn, w, metric, dom_size, param; samples) +
-            weigths_bias(w)
+            loss(solutions, non_sltns, icn, w, _metric, dom_size, param; samples) + _bias(w)
 
     _icn_ga = GA(;
         populationSize=pop_size,
@@ -74,6 +76,7 @@ function optimize!(
     metric,
     pop_size;
     sampler=nothing,
+    memoize=true,
 )
     results = Dictionary{BitVector,Int}()
     aux_results = Vector{BitVector}(undef, global_iter)
@@ -94,6 +97,7 @@ function optimize!(
             pop_size,
             iter;
             samples,
+            memoize,
         )
         aux_results[i] = weigths(aux_icn)
     end

### 2021-09-13 src/learn.jl

@@ -15,6 +15,7 @@ function learn_compose(
     metric=:hamming,
     pop_size=400,
     sampler=nothing,
+    memoize=true,
 )
     icn = ICN(; param=!isnothing(param))
     _, weigths = optimize!(
@@ -28,6 +29,7 @@ function learn_compose(
         metric,
         pop_size;
         sampler,
+        memoize
     )
     compositions = Dictionary{Composition,Int}()
     for (bv, occurences) in pairs(weigths)
@@ -68,6 +70,7 @@ function explore_learn_compose(
     configurations=explore(
         domains, concept, param; search, complete_search_limit, solutions_limit
     ),
+    memoize=true,
 )
     dom_size = maximum(domain_size, domains)
     solutions, non_sltns = configurations
@@ -81,6 +84,7 @@ function explore_learn_compose(
         metric,
         pop_size,
         sampler,
+        memoize,
     )
 end
 
@@ -119,6 +123,7 @@ function compose_to_file!(
     search_limit=1000,
     solutions_limit=100,
     sampler=nothing,
+    memoize=true,
 )
     compo, icn, _ = explore_learn_compose(
         domains,
@@ -133,6 +138,7 @@ function compose_to_file!(
         search_limit,
         solutions_limit,
         sampler,
+        memoize,
     )
 
     composition_to_file!(compo, path, name, language)

### 2021-09-13 xp/scripts/icn.jl

@@ -64,6 +64,7 @@ function icn_benchmark_unit(params)
             pop=params[:population],
             sampling=params[:sampling],
             search=string(params[:search]),
+            mem=params[:memoize],
         ),
         "json",
     )
@@ -86,8 +87,8 @@ function icn_benchmark_unit(params)
         metric = params[:metric]
         domain_size = params[:domains_size]
         domains = fill(domain(1:domain_size), domain_size)
-        func_name = "icn" * string(constraint_concept)[8:end] * "_" * string(metric)
-        func_path = datadir("compositions", func_name * ".jl")
+        # func_name = "icn" * string(constraint_concept)[8:end] * "_" * string(metric)
+        # func_path = datadir("compositions", func_name * ".jl")
 
         # Time the data retrieval/generation
         t = @timed search_space(
@@ -110,6 +111,7 @@ function icn_benchmark_unit(params)
             pop_size=params[:population],
             configurations=(solutions, non_sltns),
             sampler=params[:loss_sampler],
+            memoize=params[:memoize],
         )
         _, _, all_compos = bench.value
 

### 2021-09-13 xp/src/constants.jl

@@ -51,6 +51,8 @@ const DEFAULT_LOSS_SAMPLER = [
     nothing, root(2), root(3), constant(DEFAULT_LOSS_SAMPLING_THRESHOLD)
 ]
 
+const DEFAULT_MEMOIZATION = [true, false]
+
 # All parameters
 const ALL_PARAMETERS = Dict(
     # Search parameters
@@ -69,6 +71,7 @@ const ALL_PARAMETERS = Dict(
     :population => DEFAULT_POPULATIONS,
     :loss_sampler => DEFAULT_LOSS_SAMPLER,
     :loss_sampling_threshold => DEFAULT_LOSS_SAMPLING_THRESHOLD,
+    :memoize => DEFAULT_MEMOIZATION,
 )
 
 # ALL_PARAMETERS[:search] = [:flexible]

### 2021-09-14 xp/intro.jl
Split ICNBenchmark content

@@ -1,16 +1,0 @@
-using DrWatson
-@quickactivate "ICN benchmarks"
-
-println(
-"""
-Currently active project is: $(projectname())
-
-Path of active project: $(projectdir())
-
-Have fun with your new project!
-
-You can help us improve DrWatson by opening
-issues on GitHub, submitting feature requests,
-or even opening your own Pull Requests!
-"""
-)

### 2021-09-14 xp/scripts/icn.jl

@@ -1,175 +1,0 @@
-using Pkg
-#Pkg.add("DrWatson")
-
-# Load DrWatson (scientific project manager)
-using DrWatson
-
-# Activate the ICNBenchmarks project
-@quickactivate "ICNBenchmarks"
-
-# Pkg.instantiate()
-# Pkg.update()
-
-# Load common code to all script in ICNBenchmarks
-using ICNBenchmarks
-
-# Load other packages
-using BenchmarkTools
-using CompositionalNetworks
-using ConstraintDomains
-using Constraints
-using JSON
-
-function icn_benchmark_unit(params)
-    @info "Running a benchmark unit with" params
-
-    search_space_size = params[:domains_size]^params[:domains_size]
-    if params[:search] == :complete
-        if search_space_size > params[:complete_search_limit]
-            @warn "Unit benchmark aborted (complete) search space is too large" search_space_size params[:complete_search_limit]
-            return nothing
-        end
-        if search_space_size > params[:loss_sampling_threshold]
-            if isnothing(params[:loss_sampler])
-                @warn "Unit benchmark aborted (complete) search space is too large, and loss function is deterministically evaluated" search_space_size params[:loss_sampling_threshold] params[:loss_sampler]
-                return nothing
-            end
-        end
-    end
-    if params[:search] == :partial
-        if search_space_size < params[:partial_search_limit]
-            @warn "Unit benchmark aborted (partial) search space is too small" search_space_size params[:partial_search_limit]
-            return nothing
-        end
-        if search_space_size ≤ params[:loss_sampling_threshold]
-            if !isnothing(params[:loss_sampler])
-                @warn "Unit benchmark aborted (complete) search space is too small, and loss function is stochastically evaluated" search_space_size params[:loss_sampling_threshold] params[:loss_sampler]
-                return nothing
-            end
-        end
-    end
-    json_name = savename(
-        (
-            con=string(params[:concept][1]),
-            par=params[:concept][2],
-            csl=params[:complete_search_limit],
-            dom=params[:domains_size],
-            gen=params[:generations],
-            iter=params[:icn_iterations],
-            lang=string(params[:language]),
-            lst=params[:loss_sampling_threshold],
-            ls=string(params[:loss_sampler]),
-            metric=string(params[:metric]),
-            psl=params[:partial_search_limit],
-            pop=params[:population],
-            sampling=params[:sampling],
-            search=string(params[:search]),
-            mem=params[:memoize],
[…]

### 2021-09-14 xp/src/ICNBenchmarks.jl

@@ -1,30 +1,0 @@
-module ICNBenchmarks
-
-# usings
-using BenchmarkTools
-using CompositionalNetworks
-using ConstraintDomains
-using Constraints
-using CSV
-using DrWatson
-using Tables
-
-# imports
-import Constraints: make_error
-
-# constants
-export ALL_PARAMETERS
-export BENCHED_CONSTRAINTS
-
-# structures
-export RandomParameter, RP1 # randome parameter of length 1
-
-# others
-export search_space
-
-# includes
-include("constants.jl")
-include("search_space.jl")
-include("extra_constraints.jl")
-
-end

### 2021-09-14 xp/src/constants.jl

@@ -1,78 +1,0 @@
-# Constraints parameters
-const BENCHED_CONSTRAINTS = deepcopy(usual_constraints)
-
-const DEFAULT_CONCEPTS = [(:all_different, nothing)]
-const DEFAULT_LANGUAGES = [:Julia] # [:Julia, :C, :CPP]
-const DEFAULT_METRICS = [:hamming, :manhattan]
-const DEFAULT_COMPLETE_SEARCH_LIMIT = 6^6 # could be improved with multithreading
-const DEFAULT_PARTIAL_SEARCH_LIMIT = 4^4
-const DEFAULT_SAMPLINGS = [100, 1000]
-
-const MINIMUM_DOMAIN_SIZE = 3
-const MAXIMUM_DOMAIN_SIZE = 8 # NOTE - Check if it is too high
-
-function domains_sizes(;
-    min_domain_size=MINIMUM_DOMAIN_SIZE, max_domain_size=MAXIMUM_DOMAIN_SIZE
-)
-    return collect(min_domain_size:max_domain_size)
-end
-
-# Genetic Algorithms default parameters
-const MAXIMUM_TOTAL_ITERATIONS = 12 # order => 2^15 = 32768
-const MINIMUM_ICN_ITERATIONS = 4 # order => 2^5= 32
-const MINIMUM_GENERATIONS = 4 # order => 2^5 = 32
-
-function icn_iterations(;
-    min_icn_iterations=MINIMUM_ICN_ITERATIONS,
-    min_generations=MINIMUM_GENERATIONS,
-    max_total_iterations=MAXIMUM_TOTAL_ITERATIONS,
-)
-    maximum_icn_iterations = max_total_iterations - min_generations
-    return [2^i for i in min_icn_iterations:2:maximum_icn_iterations]
-end
-
-function generations(;
-    min_icn_iterations=MINIMUM_ICN_ITERATIONS,
-    min_generations=MINIMUM_GENERATIONS,
-    max_total_iterations=MAXIMUM_TOTAL_ITERATIONS,
-)
-    maximum_generations = max_total_iterations - min_icn_iterations
-    return [2^i for i in min_generations:2:maximum_generations]
-end
-
-const DEFAULT_POPULATIONS = [2^i for i in 6:2:10] # 2^5=32 to 2^10 = 1024
-
-const DEFAULT_LOSS_SAMPLING_THRESHOLD = 2^10
-
-root(n) = x -> round(Int, x^(1 / n))
[…]

### 2021-09-14 xp/src/extra_constraints.jl

@@ -1,23 +1,0 @@
-# This files stores constraints or concepts that are not yet available within usual_constraints
-
-# Sample for all_different
-# const all_different = Constraint(
-#     concept = concept_all_different,
-#     error = make_error(:all_different),
-#     syms = Set([:permutable]),
-# )
-
-## no overlap
-function concept_no_overlap(x, l)
-    for i in 1:length(x), j in i+1:length(x)-1
-        x[i]+l[i] > x[j] && x[j]+l[j] > x[i] && return false
-    end
-    return true
-end
-
-const no_overlap = Constraint(
-    concept = concept_no_overlap,
-    error = make_error(:no_overlap),
-)
-
-push!(BENCHED_CONSTRAINTS, :no_overlap => no_overlap)

### 2021-09-14 xp/src/search_space.jl

@@ -1,75 +1,0 @@
-
-function search_space(
-    dom_size,
-    concept,
-    param=nothing;
-    search=:flexible,
-    complete_search_limit=1000,
-    solutions_limit=100,
-)
-    # Define the domains by on the domain size
-    domains = fill(domain(1:dom_size), dom_size)
-
-    # # Determine if the search is partial or complete
-    if search == :flexible
-        search = sum(domain_size, domains) < complete_search_limit ? :complete : :partial
-    end
-
-    # Define the output folder and make the related path if necessary
-    output_folder = joinpath(datadir("search_spaces"))
-    mkpath(output_folder)
-
-    # Define the file names TODO: make better metaprogramming
-    name_solutions = ""
-    name_non_sltns = ""
-    if search == :complete
-        configurations = :solutions
-        name_solutions = savename(
-            @dict constraint = Symbol(concept) param dom_size search configurations
-        )
-        configurations = :non_sltns
-        name_non_sltns = savename(
-            @dict constraint = Symbol(concept) param dom_size search configurations
-        )
-    else
-        configurations = :solutions
-        name_solutions = savename(
-            @dict constraint = Symbol(concept) param dom_size search solutions_limit configurations
-        )
-        configurations = :non_sltns
-        name_non_sltns = savename(
-            @dict constraint = Symbol(concept) param dom_size search solutions_limit configurations
-        )
-    end
-    file_solutions = joinpath(output_folder, "$name_solutions.csv")
-    file_non_sltns = joinpath(output_folder, "$name_non_sltns.csv")
-
-    # Check if existing data are present
-    has_data = isfile(file_solutions) && isfile(file_non_sltns)
-
-    function read_csv_as_set(file)
-        configs = Set{Vector{Int}}()
-        for r in CSV.Rows(file; header=false, type=Int)
-            push!(configs, collect(Int, r))
-        end
-        return configs
-    end
-
-    # Load or compute the exploration of the search space
-    solutions, non_sltns = if has_data
-        read_csv_as_set(file_solutions), read_csv_as_set(file_non_sltns)
-    else
-        explore(domains, concept, param; search, solutions_limit)
-    end
-
-    # Save the data locally if generated on this run
-    if !has_data
-        files = [file_solutions, file_non_sltns]
-        configs = [solutions, non_sltns]
-        for (file, config) in zip(files, configs), x in config
-            CSV.write(file, Tables.table(reshape(x, (1, length(x)))); append=true)
-        end
-    end
-
-    return solutions, non_sltns, has_data
-end

### 2021-09-14 xp_old/compositions/icn_all_diff_hamming.jl

@@ -1,12 +1,0 @@
-function icn_all_diff_hamming(x; X = zeros(length(x), 1), param=nothing, dom_size)
-    CompositionalNetworks.tr_in(Tuple([CompositionalNetworks.tr_count_eq_left]), X, x, param)
-    for i in 1:length(x)
-        X[i,1] = CompositionalNetworks.ar_sum(@view X[i,:])
-    end
-    return CompositionalNetworks.ag_count_positive(@view X[:, 1]) |> (y -> CompositionalNetworks.co_identity(y; param, dom_size, nvars=length(x)))
-end
-#Generation:
- #8.348 s (53084802 allocations: 2.95 GiB)
-#Use:
-# 1.052 μs (6 allocations: 352.0 bytes)
-# test

### 2021-09-14 xp_old/concepts.jl

@@ -1,42 +1,0 @@
-# all_diff
-concept_all_diff(x) = allunique(x)
-
-#dist_diff
-function concept_dist_diff(x)
-    for i in 1:length(x)-3
-        if abs(x[i] - x[i+1]) == abs(x[i+2] - x[i+3])
-            return false
-        end
-    end
-    return true
-end
-
-#equals
-concept_all_eq(x) = all(y -> y == x[1], x[2:end])
-
-#less_than_param
-concept_less_than_param(x, param) = x[1] ≤ param
-
-
-#ordered
-concept_ordered(x) = issorted(x)
-
-#no_overlap (general idea, still need to review it)
-function concept_no_overlap(x, l)
-    for i in 1:length(x)
-        for j in i+1:length(x)-1
-            if (x[i]+l[i] > x[j] && x[j]+l[j] > x[i])
-                return false
-            end
-        end
-    end
-    return true
-end
-
-
-#sum of vars = c
-concept_sum_equal_param(x; param) = sum(x) == param
-
-
-#concept_list = [concept_all_diff]
-concept_list = [concept_all_diff, concept_dist_diff, concept_all_eq, concept_less_than_param, concept_ordered]

### 2021-09-14 xp_old/domains.jl

@@ -1,4 +1,0 @@
-# 4 variables in 1:5 (search space 5^4 = 625)
-#dom = [1,2,3,4,5]
-dom = [1,2,3,4]
-domains = [domain(dom) for i in 1:4]

### 2021-09-14 xp_old/helpers.jl

@@ -1,133 +1,0 @@
-"""
-Contains all icn functions that somehow aren't recognized by the compiler
-TODO:: Fix it later
-"""
-
-
-function as_int(v::AbstractVector)
-    n = 0
-    for (i, b) in enumerate(v)
-        n += b ? 2^(i - 1) : 0
-    end
-    return n
-end
-
-nbits(icn) = mapreduce(l -> exclu(l) ? nbits_exclu(l) : length(l), +, layers(icn))
-
-functions(layer) = layer.functions
-
-symbol(layer, i) = collect(keys(functions(layer)))[i]
-
-nbits_exclu(layer) = ceil(Int, log2(length(layer)))
-
-exclu(layer) = layer.exclusive
-
-layers(icn) = [icn.transformation, icn.arithmetic, icn.aggregation, icn.comparison]
-
-weigths(icn) = icn.weigths
-
-is_viable(layer, w) = exclu(layer) ? as_int(w) < length(layer) : any(w)
-
-function is_viable(icn::ICN, weigths)
-    _start = 0
-    _end = 0
-
-    for layer in layers(icn)
-        _start = _end + 1
-        _end += exclu(layer) ? nbits_exclu(layer) : length(layer)
-
-        w = @view weigths[_start:_end]
-
-        !is_viable(layer, w) && return false
-    end
-    return true
-end
-is_viable(icn::ICN) = is_viable(icn, weigths(icn))
-
-function weights!(icn, weigths)
-    @assert length(weigths) == nbits(icn)
-    icn.weigths = weigths
-end
-
-@unroll function tr_in(tr, X, x, param)
-    @unroll for i in 1:length(tr)
-        X[:,i] = tr[i](x; param)
-    end
-end
-
-function _compose(icn::ICN)
-    !is_viable(icn) && (return ((x; param=nothing, dom_size=0) -> typemax(Float64)), [])
-
-    funcs = Vector{Vector{Function}}()
-    symbols = Vector{Vector{Symbol}}()
-
-    _start = 0
-    _end = 0
-
-    for layer in layers(icn)
-        _start = _end + 1
-        _end += exclu(layer) ? nbits_exclu(layer) : length(layer)
-
-        if exclu(layer)
-            f_id = as_int(@view weigths(icn)[_start:_end])
-            s = symbol(layer, f_id + 1)
-            push!(funcs, [functions(layer)[s]])
-            push!(symbols, [s])
-        else
-            layer_funcs = Vector{Function}()
-            layer_symbs = Vector{Symbol}()
[…]

### 2021-09-14 xp_old/params.jl

@@ -1,4 +1,0 @@
-include("../src/metrics.jl")
-
-metrics = [hamming, manhattan]
-#metrics = [hamming]

### 2021-09-14 xp_old/xp.jl

@@ -1,89 +1,0 @@
-using Unrolled
-using ConstraintDomains
-using CompositionalNetworks
-using BenchmarkTools
-
-using Distributed
-
-# if nworkers == 1
-#     addprocs()
-# end
-
-# println("N-workers: $(nworkers())")
-n_threads = Threads.nthreads()
-println("N-THREADS: $n_threads")
-
-include("domains.jl")
-include("concepts.jl")
-include("params.jl")
-include("helpers.jl")
-
-
-
-if (isdir("xp"))
-    cd("xp")
-    cd("compositions")
-elseif (contains(pwd()[end-20:end], "CompositionalNetworks"))
-    mkdir("xp")
-    cd("xp")
-    mkdir("compositions")
-    cd("compositions")
-end
-
-
-function write_benchmarks(path, data)
-    file = open(path, "a")
-    write(file, data)
-    close(file)
-end
-
-C = 640
-global_iters = n_threads
-local_iters = C ÷ global_iters
-
-#for concept in concept_list
-#    for metric in metrics
-
-concept = concept_list[1]
-metric = metrics[2]
-        m = metric
-        println("$concept-$metric")
-        func_name = "icn$(String(Symbol(concept))[8:end])_$(String(Symbol(metric)))"
-        # TODO: Fix param value
-        param = length(iterate(methods(concept))[1].sig.parameters) == 2 ? nothing : rand(dom)
-        path = "$(func_name)_$n_threads.jl"
-        icn = ICN()
-        g_b = @benchmark ($icn = compose_to_file!($concept, "$($func_name)", $path,
-                         domains=domains, param=$param, global_iter=global_iters,local_iter=local_iters ,  metric=$metric)) samples = 2
-
-        g_b_time = BenchmarkTools.prettytime(time(g_b))
-        g_b_memory = BenchmarkTools.prettymemory(memory(g_b))
-        g_b_allocs = allocs(g_b)
-
-        write_benchmarks(path, "#Finding composition:\n#$g_b_time ($g_b_allocs allocation$(g_b_allocs == 1 ? "" : "s"): $g_b_memory)")
-
-        X_sols, X = complete_search_space(domains, concept, param)
-        #loss_value = fitness(icn.weigths)
-        loss_value = fitness(weigths(icn), X, X_sols, icn, metric, param)
-        include("compositions/$path")
[…]

### 2021-09-29 paper/paper.tex
Paper Wednesday, by Jeff

@@
 \setcounter{page}{1}
+
+\usepackage{xspace}
+
+\usepackage{subcaption}
+
+\usepackage{hyperref}
+
+\hypersetup{colorlinks=true,linkcolor=blue,urlcolor=blue}
+
+\newcommand{\ie}{\emph{i.e.}}
+\newcommand{\cp}{\textsc{CP}\xspace}
+\newcommand{\csp}{\textsc{CSP}\xspace}
+\newcommand{\cop}{\textsc{COP}\xspace}
+\newcommand{\efsp}{\textsc{EFSP}\xspace}
+\newcommand{\efop}{\textsc{EFOP}\xspace}
+\newcommand{\cfn}{\textsc{CFN}\xspace}
+\newcommand{\wcsp}{\textsc{WCSP}\xspace}
+\newcommand{\cbls}{\textsc{CBLS}\xspace}
+\newcommand{\ghost}{\textsc{GHOST}\xspace}
+\newcommand{\cppn}{\textsc{CPPN}\xspace}
+\newcommand{\icn}{\textsc{ICN}\xspace}
+
+\newcommand{\jc}{\href{https://github.com/JuliaConstraints}{JuliaConstraints}\xspace}
+\newcommand{\cdjl}{\href{https://github.com/JuliaConstraints/ConstraintDomains.jl}{ConstraintDomains.jl}\xspace}
+\newcommand{\cnjl}{\href{https://github.com/JuliaConstraints/CompositionalNetworks.jl}{CompositionalNetworks.jl}\xspace}
+\newcommand{\cjl}{\href{https://github.com/JuliaConstraints/Constraints.jl}{Constraints.jl}\xspace}
+\newcommand{\cmjl}{\href{https://github.com/JuliaConstraints/ConstraintModels.jl}{ConstraintModels.jl}\xspace}
+\newcommand{\lssjl}{\href{https://github.com/JuliaConstraints/LocalSearchSolvers.jl}{LocalSearchSolvers.jl}\xspace}
+\newcommand{\cblsjl}{\href{https://github.com/JuliaConstraints/CBLS.jl}{CBLS.jl}\xspace}
+
+\newcommand{\cproblem}[3]%
+{\begin{trivlist}
+  \item[]%
+    \textbf{Problem:} \textsc{#1}\\
+    \textit{Input:} #2\\
+    \textit{Question:} #3
+  \end{trivlist}%
+}
 
@@
 \begin{abstract}
-
-Interpretable Compositional Networks (ICN), a variant of neural networks, that allows the user to get interpretable results, unlike regular artificial neural networks.
-An ICN is a glass-box producing functions composition that scale with the size of the input, allowing a learning phase on relatively small spaces.
-This presentation covers the different Julia packages and paradigms involved, a set of use-case, current limitations, future developments, and hopefully possible collaborations.
-
+  Interpretable Compositional Networks (ICNs) are a neural network variant for combinatorial function learning that allows the user to obtain interpretable results, unlike ordinary artificial neural networks.
+An ICN outputs a composition of functions that scale with the size of the input, allowing a learning phase on relatively small spaces.
+  \cnjl is a pure Julia package that exploits the language's meta-programming, parallelism and multiple dispatch features to produce learned compositions in mathematical and programming languages such as Julia, C or C++.
 \end{abstract}
 
+\section{Introduction}
+\label{sec:introduction}
+
+The discipline of combinatorial optimization consists in finding an optimal configuration of elements within a finite set.
+Such a set is usually subject to constraints that can be represented by functions that are often highly combinatorial.
+These constraints are mostly formulated as concepts, boolean functions indicating whether each constraint is respected or not.
+A solution (sometimes called a satisfactory solution, depending on the domain) is a configuration that respects all the constraints.
+
+Different domains such as operational research, constraint programming, metaheuristics, propose methods to find (satisfactory or optimal) solutions.
+Among these methods, some can or could benefit from a finer granularity in the evaluation of the impact of each constraint on a given configuration.
+
+In \cite{richoux2020automatic}, we introduced Interpretable Compositional Networks (ICNs), a neural network variant for combinatorial function learning that allows the user to obtain interpretable results, unlike ordinary artificial neural networks.
+An ICN outputs a composition of functions that scale with the size of the input, allowing a learning phase on relatively small spaces.
+
+This work, consists in a library that implements ICNs in the Julia programming language \cite{bezanson2017julia}.
+Although we present a direct application of ICNs in the following subsection, this neural network framework is simple to extend to learn other combinatorial and non-combinatorial functions.
+
+\subsection{Application to Constraint Programming}
+\label{subsec:cpcontext}
+
+Constraint Satisfaction Problem (\csp) and Constrained Optimization Problem (\cop) are constraint-based problems where constraints can be seen as predicates (\emph{concepts}) allowing or forbidding some combinations of variable assignments.
+Such a formulation corresponds well to the so-called complete solution methods which guarantee an exploration of all solutions.
+Unfortunately, due to the highly combinatorial nature of some problems, these methods cannot always converge in a reasonable time.
+
+\begin{figure}[t]
+  \centerline{\includegraphics[width=5cm]{figs/juliaconstraints.png}}
+  \caption{Logo of the \jc organization on GitHub that hosts, among other things, the \cnjl package.}
+  \label{fig:juliaconstraints}
+\end{figure}
+
+On the other hand, Constraint-Based Local Search (\cbls) is a family of metaheuristics in which the neighborhood is constructed on the basis of an \emph{error function}, itself a quantitative representation of how far the current configuration is from an admissible solution.
+We refer to the corresponding problems as Error Function Satisfaction Problem (\efsp) and Error Function Optimization Problem (\efop).
+In the case of \cbls, this is computed as a function of the constraints which are not currently satisfied.
+Error functions may be derived from the constraint satisfaction problem structure, hand-coded, automatically acquired by some machine learning process, or constructed as a combination of these methods.
+It should be noted that, the best performing systems resort to hand-tuned error functions, as witness~\cite{DBLP:books/sp/18/CodognetMDA18}.
+
+As illustrated by Figure~\ref{fig:landscape}, a well-designed error function helps in converging faster towards better-quality solutions.
+It does, however, introduce an additional layer of complexity to the user.
+
+\begin{figure}[t]
+  \centering
+  \subcaptionbox{CSP landscape\label{sfig:csp}}{
+    \includegraphics[width = .95\columnwidth]{figs/csp_landscape_complex_zero_big.png}
+  }
+
+  \subcaptionbox{EFSP landscape\label{sfig:efsp}}{
+    \includegraphics[width = .95\columnwidth]{figs/efsp_landscape_complex_zero_big.png}
+  }\caption{
+    Comparison of a Constraint Satisfaction Problem (CSP) and an Error
+    Function Satisfaction Problem (\efsp) landscapes. The finer scale
+    in the error heuristic of the \efsp leads to a better convergence
+    rate.}\label{fig:landscape}
+\end{figure}
+
+\subsection{Interpretable Compositional Networks (ICN)}
+\label{subsec:icn}
+
+In \cite{richoux2020automatic}, we proposed a neural network inspired by Compositional Pattern-Producing Networks (\cppn) to learn (highly) combinatorial functions as non-linear combinations of elementary operations.
+\(\cppn\)s~\cite{stanley2007cppn} are themselves a variant of artificial neural networks.
+While neurons in regular neural networks usually  contain sigmoid-like functions only (such  as ReLU, \ie{} Rectified Linear  Unit), \cppn's neurons can contain many other kinds of functions: sigmoids, Gaussians, trigonometric functions, and linear functions among others.
+\(\cppn\)s are often used to generate 2D or 3D images by applying the function modeled by a \cppn giving each pixel individually as input, instead of considering all pixels at once.
+This simple trick allows the learned \cppn model to produce images of any resolution.
+
+We propose our variant by taking these two principles from \cppn: having neurons containing one operation among many possible ones, and handling inputs in a size-independent fashion.
+Due to their interpretable nature, we named our variant \textbf{Interpretable Compositional Networks} (\icn).
+
+Although ICNs are not limited to learning function for \efsp/\efop, the original structure was designed to learn compositions weighted in accordance to the hamming metric \cite{richoux2020automatic}.
+The hamming distance between a configuration and its closest solution is a meaningful indicator of the number of variables needed to be changed.
+Manhattan and other variants of minkowski comparison are other example of possible metrics.
+
+An ICN is made of several layers of (possibly exclusive) operations such that the first layers accept a vector as input and the last layer returns a single numerical value.
+Generally, layers alter there input in three possible ways: \emph{increment}, \emph{decrement}, or \emph{conservation} of the dimension of the input.
+
+\begin{figure}[t]
+  \centerline{\includegraphics[width=\columnwidth]{figs/model_nn.png}}
+  \caption{Scheme of the basic 4-layers ICN model used in \efsp/\efop.
+Layers with blue neurons have mutually exclusive operations.
+the \emph{transformation} layer increases the input dimension, the \emph{arithmetic} and \emph{aggregation} layers reduces it, and the \emph{comparison} layer leaves it untouched.
+The figure is taken from \cite{richoux2020automatic}.}
+  \label{fig:model_icn}
+\end{figure}
+
+In the context of \efsp/\efop, the output should be non-negative if the constraint is violated and equal to 0 otherwise. As shown by Figure \ref{fig:model_icn}, our \(\icn\)s are
+composed of four layers, each of them having a specific purpose and
+themselves composed of neurons applying  a unique operation each.  All
+neurons  from  a  layer  are  linked to  all  neurons  from  the  next
+layer. The weight on each link is purely binary: its value is either 0
+or 1. This restriction is  crucial to obtain interpretable functions. We refer the reader to \cite{richoux2020automatic} for a more comprehensive definition.
+
+\subsection{A first C++ library}
+\label{subsec:cplusplus}
+
+In \cite{richoux2020automatic} we introduced the concept of ICN and a first implementation for Constraint Programming as a C++ library\footnote{\url{https://github.com/richoux/LearningUtilityFunctions}}.
+The results were evaluated through the GHOST C++ solver \cite{richoux2016ghost} and serve as a proof of concept that most of the models give scalable functions, and remain fairly effective using incomplete training sets.
+
+As mentioned in Section \ref{sec:xlc}, \cnjl generates code for a direct use in Julia, but also exports usable code for both the GHOST(C++) and AdaptiveSearch (C)\footnote{This feature is a WIP at the moment this article is submitted, but is expected to be completed in the coming weeks.}.
+The code exported in C++, that is the \emph{operations} of each neuron, is strongly inspired by our C++ library.
+
+\section{The example of the \jc framework}
+\label{sec:juliaconstraints}
+
+The \cnjl package is part of the \jc GitHub organization that proposes a first collection of Julia packages for Constraint-Based Local Search (CBLS), a subfield of Constraint Programming.
+
+The main goal of this framework is to provide a set of high level semi-automatic tools which strike a good compromise between efficiency and ease of modeling.
+
+\subsection{History of \jc}
+\label{subsec:history}
+
+The LocalSearchSolvers.jl package is a Constraint-Based Local Search framework started in Fall 2020 and inspired by other CBLS solvers such as GHOST (C++) and AdaptiveSearch (C) that allows users to tune their own solver.
+
+During the development of LocalSearchSolvers.jl, we decided to split the code and functionality of the original framework into several independent packages for ease of maintenance and in hopes of providing common tools for other constraint programming packages in Julia.
+
+During 2021, these tools were collected into the \jc ecosystem as follows
 \begin{itemize}
-    \item introduction to
-    \begin{itemize}
-        \item CP
-        \item Metaheuristics
-        \item CBLS and EFN
-    \end{itemize}
-    \item Pros of doing things in Julia
-    \item paper organization/content
+  \item \cdjl: creation of discrete, continuous, and arbitrary domain (beta status)
+  \item \cjl: generation of constraints from a boolean concept or an error function. Also list usual constraints and their properties (beta status)
+  \item \cnjl: stable
+  \item \lssjl: A CBLS framework in Julia with built-in parallel and distributed scalability
+  \item \cblsjl: A MOI/JuMP wrapper for LocalSearchSolvers
+  \item \cmjl: list of CP models for \cblsjl and \lssjl
 \end{itemize}
 
-\section{Introduction}
+\begin{figure*}[t]
+  \centerline{\includegraphics[page=1, width=.9\textwidth]{figs/overview.pdf}}
+  \caption{Overview of the Constraint Programming ecosystem in Julia, including \jc.
+\emph{Front-End} and \emph{Internals} refer to the latest.
+Note that, at the moment of the writing, \texttt{ConstraintProgrammingExtension.jl}, \texttt{JuCP.jl}, and \texttt{COPInstances.jl} are temporary names.}
+  \label{fig:overview}
+\end{figure*}
 
-The \LaTeX{} document preparation system is a special version of the
-\TeX; typesetting program where-in a collection of \TeX; commands
-are added to \LaTeX{} to simplify typesetting. Importantly, it allows
-the author to concentrate on the logical structure of the document
-rather than its visual layout.\vskip 6pt
-Moreover, \LaTeX{} provides a consistent and comprehensive document
-preparation interface. There are simple-to-use commands for
-generating a table of contents, lists of figures and/or tables, and indexes.
-\LaTeX{} can automatically number list entries, equations, figures,
-tables, and footnotes, as well as articles, sections, and subsections.
-Using this numbering system, bibliographic citations, page
-references, and cross references to any other numbered entity (e.g.
-article, section, equation, figure, list entry, etc.) become quite simple
-and straightforward. The use of \LaTeX{} document classes allows
-a simple change of class to transform the appearance of your document.\vskip 6pt
-\LaTeX{} is a powerful tool for managing long and complex documents.
-In particular, partial processing enables long documents to
-be produced article by article without losing sequential information.
-The use of document classes allows a simple change of style
-(or style option) to transform the appearance of your document.
+The global state of the Constraint Programming ecosystem is presented in Figure \ref{fig:overview}.
+Beside the CBLS framework of \jc, other solvers are complete search methods, such as ConstraintSolver.jl\footnote{\url{https://github.com/Wikunia/ConstraintSolver.jl}}, CPLEXCP.jl\footnote{\url{https://github.com/dourouc05/CPLEXCP.jl}}, and SeaPearl.jl \cite{chalumeau2021seapearl}.
 
-\section{The JuliaCon Article Class}
-\label{sec:documentclass}
-%
-The juliacon class file preserves the standard LATEX{} interface such
-that any document that can be produced using the standard LATEX{}
-article class can also be produced with the class file.\vskip 6pt
-It is likely that the make up will change after file submission. For
-this reason, we ask you to ignore details such as slightly long lines,
-page stretching, or figures falling out of synchronization, as these
-details can be dealt with at a later stage.\vskip 6pt
-Use should be made of symbolic references (\verb|\ref| ) in order to
-protect against late changes of order, etc.
+\subsection{A simple example with \texttt{CBLS.jl}}
+\label{subsec:example}
 
-\section{USING THE JuliaCon Article CLASS FILE}
+A Magic Square of order $n$ is composed of $n^2$ distinct values ranging from $1$ to $n^2$ layed out as a square array $X$.
+These values need to be arranged such that the sums of each diagonal, row and column be equal to the same value, the magic sum $\Sigma$.
 
-If the file \verb|juliacon.cls|  is not already in the appropriate system directory
-for \LaTeX{} files, either arrange for it to be put there or copy
-it to your working directory. The \verb|juliacon|  document class is implemented
-as a complete class, not a document style option. In order to
-use the \verb|juliacon  document class, replace \verb|article|  by \verb|juliacon|  in the
-\verb|\documentclass|  command at the beginning of your document:
-\vskip 6pt
-\begin{centering}
-    \verb|\documentclass{article}|  \end{centering}
-\vskip 6pt
-replace by
-\vskip 6pt
- \verb|\documentclass{juliacon}|  \vskip 6pt
-In general, the following standard document \verb|style|  options should
-{ \itshape not} be used with the {\footnotesize \itshape article} class file:
-\begin{enumerate}
-\item[(1)] \verb|10pt|,  \verb|11pt|,  \verb|12pt|   ? unavailable;
-\item[(2)] \verb|twoside|  (no associated style file) ? \verb|twoside|  is the default;
-\item[(3)] \verb|fleqn|, \verb|leqno|, \verb|titlepage| ? should not be used;
-\end{enumerate}
+\[ \sum_{j=1}^{n} X_{ij} = \sum_{i=1}^{n} X_{ij} = \sum_{j=1}^{n} X_{ii} = \sum_{j=1}^{n} X_{i,n-i+1} = \frac{n(n^{2}+1)}{2} = \Sigma \]
 
-\section{Additional Document Style Options}
-\label{sec:additional_doc}
-%
-The following additional style option is available with the \verb|juliacon|  class file:
-\vskip 6pt
-Please place any additional command definitions at the very start of
-the \LaTeX{} file, before the \verb|\begin{document}|. For example, user-defined
-\verb|\def|  and \verb|\newcommand|   commands that define macros for
-technical expressions should be placed here. Other author-defined
-macros should be kept to a minimum.
-\vskip 6pt
-Commands that differ from the standard \LaTeX{} interface, or that
-are provided in addition to the standard interface, are explained in
-this guide. This guide is not a substitute for the \LaTeX{} manual itself.
-Authors planning to submit their papers in \LaTeX{} are advised to use
-\verb|juliacon.cls|  as early as possible in the creation of their files.
+\begin{figure}[t]
+  \centerline{\includegraphics[page=1]{figs/magicsquare.pdf}}
+  \caption{A solved magic-square of size $3$. The \emph{magic sum} is equal to $15$. Image from Wikipedia's user \textsc{Phidauex}.}
+  \label{fig:magicsquare}
+\end{figure}
 
-%
-%
-%
-%
-\begin{table*}[t]
-\tabcolsep22pt
-\tbl{If necessary, the tables can be extended both columns.}{
-\begin{tabular}{|l|l|c|c|}\hline
-Label & \multicolumn{1}{c|}{Description}
-& Number of Users &
-Number of Queries\\\hline
-Test 1 & Training Data &
-\smash{\raise-7pt\hbox{70}} & 104\\
-\cline{1-2}\cline{4-4}
-Test 2 & Testing Data I & & 105\\\hline
-Test 3 & Testing Data II & 30 & 119\\\hline
-& Total & 100 & 328\\\hline
-\end{tabular}}
-\label{tab:symbols}
-\begin{tabnote}
-This is an example of table footnote.
-\end{tabnote}
-\end{table*}
-% \begin{figure*}[t]
-% \centerline{\includegraphics[width=11cm]{juliagraphs.png}}
-% \caption{If necessary, the images can be extended both columns.}
-%   \label{fig:sample_image}
-% \end{figure*}
+In \texttt{ConstraintModels.jl}, we use the following code to generate a magic-square instance.
 
-\section{Additional features}
-\label{sec:additional_faci}
-In addition to all the standard \LaTeX{} design elements, the \verb|juliacon|  class file includes the following features:
-In general, once you have used the additional \verb|juliacon.cls| facilities
-in your document, do not process it with a standard \LaTeX{} class
-file.
+\begin{lstlisting}[language = Julia]
+# Import CBLS.jl and JuMP.jl
+using CBLS, JuMP
 
-\subsection{Titles, Author's Name, and Affiliation}
-\label{subsub:title_auth}
-The title of the article, author's name, and affiliation are used at the
-beginning of the article (for the main title). These can be produced
-using the following code:
+function magic_square(n::Integer)
+model = Model(CBLS.Optimizer)
+N = n^2
+Σ = n * (N + 1) / 2
+@variable(model, 1 <= X[1:n, 1:n] <= N, Int)
+@constraint(model, vec(X) in AllDifferent())
+for i in 1:n
+  @constraint(model, X[i,:] in Linear(Σ))
+  @constraint(model, X[:,i] in Linear(Σ))
+end
+@constraint(model,
+  [X[i,i] for i in 1:n] in Linear(Σ)
+)
+@constraint(model,
+  [X[i,n+1-i] for i in 1:n] in Linear(Σ)
+)
+  return model, X
+end
 
-\begin{verbatim}
-\title{ This is an example of article title} }
-\author{
-   \large 1st Author \\[-3pt]
-   \normalsize 1st author's affiliation  \\[-3pt]
-    \normalsize 1st line of address \\[-3pt]
-    \normalsize 2nd line of address \\[-3pt]
-    \normalsize	1st author's email address \\[-3pt]
-  \and
-   \large 2nd Author \\[-3pt]
-   \normalsize 2nd author's affiliation  \\[-3pt]
-    \normalsize 1st line of address \\[-3pt]
-    \normalsize 2nd line of address \\[-3pt]
-    \normalsize	2nd author's email address \\[-3pt]
-\and
-   \large 3rd Author \\[-3pt]
-   \normalsize 3rd author's affiliation  \\[-3pt]
-    \normalsize 1st line of address \\[-3pt]
-    \normalsize 2nd line of address \\[-3pt]
-    \normalsize	3rd author's email address \\[-3pt]
-}
-\maketitle
-\end{verbatim}
-
-\subsection{Writing Julia code}
-
-A special environment is already defined for Julia code,
-built on top of \textit{listings} and \textit{jlcode}.
-
-\begin{verbatim}
-\begin{lstlisting}[language = Julia]
-using Plots
-
-x = -3.0:0.01:3.0
-y = rand(length(x))
-plot(x, y)
-\end{lstlisting}
-\end{verbatim}
-\begin{lstlisting}[language = Julia]
-using Plots
-
-x = -3.0:0.01:3.0
-y = rand(length(x))
-plot(x, y)
+# Create a magic square instance of size 4x4
+m, X = magic_square(4)
+# Optimize
+optimize!(m)
+# Output values
+@info values.(X)
 \end{lstlisting}
 
-\subsection{Abstracts, Key words, term etc...}
-\label{subsub:abs_key_etc}
+\section{A flexible implementation}
+\label{sec:implementation}
 
-At the beginning of your article, the title should be generated
-in the usual way using the \verb|\maketitle|  command. For genaral tem and keywords use
-\verb|\terms|,
-\verb|\keywords|  commands respectively. The abstract should be enclosed
-within an abstract environment, All these environment
-can be produced using the following code:
-\begin{verbatim}
-\terms{Experimentation, Human Factors}
+\cnjl extends the work of the first C++ library used in \cite{richoux2020automatic}.
+As mentioned in Section \ref{subsec:cplusplus}, that library used for prototyping was limited through several aspects: this, this, and that.
+Obviously, those limitations could have been lifted in the first library.
+However, we decided to use some features of the Julia language to ease provisioning flexibility and broader features to ICNs.
 
-\keywords{Face animation, image-based modelling...}
+\paragraph*{Block about Julia} test
 
-\begin{abstract}
-In this paper, we propose a new method for the
-systematic determination of the model's base of
-time varying delay system. This method based on
-the construction of the classification data related
-to the considered system. The number, the orders,
-the time delay and the parameters of the local
-models are generated automatically without any
-knowledge about the full operating range of the
-process. The parametric identification of the local
-models is realized by a new recursive algorithm for
-on line identification of systems with unknown time
-delay. The proposed algorithm allows simultaneous
-estimation of time delay and parameters of
-discrete-time systems. The effectiveness of
-the new method has been illustrated through
-simulation.
-\end{abstract}
+\subsection{Layers of operations}
+\label{subsec:layers}
 
-\end{verbatim}
+A layer is defined by a non-empty collection of (possibly mutually exclusive) operations.
+The original ICN is composed of four layers: \emph{transformation} (exclusive), \emph{arithmetic}, \emph{aggregation}, and \emph{comparison}.
 
-\section{Some guidelines}
-\label{sec:some_guide}
-The following notes may help you achieve the best effects with the
-\verb|juliacon|  class file.
+\begin{lstlisting}[language = Julia]
+struct Layer
+  functions::LittleDict{Symbol, Function}
+  exclusive::Bool
+end
+\end{lstlisting}
 
-\subsection{Sections}
-\label{subsub:sections}
-\LaTeXe{}  provides four levels of section headings and they are all
-defined in the \verb|juliacon|  class file:
+\begin{lstlisting}[language = Julia]
+# Transformation defined for a vector
+tr_identity(x; param=nothing) = identity(x)
+
+# Transformation defined for the index of a vector
+tr_count_eq(i, x; param=nothing) =
+  count(y -> x[i] == y, x) - 1
+tr_count_eq_param(i, x; param) =
+  count(y -> y == x[i] + param, x)
+
+# Generating vectorized versions
+lazy(tr_count_eq)
+lazy_param(tr_count_eq_param)
+\end{lstlisting}
+
+\begin{lstlisting}[language = Julia]
+function transformation_layer(param=false)
+  transformations = LittleDict{Symbol,Function}(
+    :identity => tr_identity,
+    :count_eq => tr_count_eq,
+    # ...
+  )
+  if param
+    tr_param = LittleDict{Symbol, Function}(
+      :count_eq_param => tr_count_eq_param,
+      # ...
+    )
+    transformations = LittleDict(
+      union(transformations, tr_param)
+    )
+  end
+  return Layer(transformations, false)
+end
+\end{lstlisting}
+
+All methods required for the definition of additional layers are all available.
+
+\subsection{Composition}
+\label{subsec:composition}
+
+We store the composition of an ICN in three forms:
 \begin{itemize}
-\item \verb|\section|
-\item \verb|\subsection|
-\item \verb|\subsubsection|
-\item \verb|\paragraph|
+  \item A collection of \texttt{Symbol}s per layer of the ICN that allows the generation of code as a string
+  \item The Julia object of type \texttt{Function} to apply the composition directly
 \end{itemize}
-Section headings are automatically converted to allcaps style.
-\subsection{Lists}
-\label{sec:lists}
-%
-The \verb|juliacon|   class file provides unnumbered lists using the
-\verb|unnumlist|  environment for example,
 
-\begin{unnumlist}
-\item First unnumbered item which has no label and is indented from the
-left margin.
-\item Second unnumbered item.
-\item Third unnumbered item.
-\end{unnumlist}
-The unnumbered list which has no label and is indented from the
-left margin. was produced by:
-\begin{verbatim}
-\begin{unnumlist}
-\item First unnumbered item...
-\item Second unnumbered item...
-\item Third unnumbered item...
-\end{unnumlist}
-\end{verbatim}
+\begin{lstlisting}[language = Julia]
+struct Composition{F<:Function}
+  code::Dict{Symbol,String}
+  f::F
+  symbols::Vector{Vector{Symbol}}
+end
+\end{lstlisting}
 
-The \verb|juliacon|  class file also provides hyphen list using the
-\verb|itemize|  environment for example,
+The result output by an ICN is a composition, a mathematical function composed of several basic operations.
+The \texttt{code} function returns the definition of a composition in either a mathematical or programming language.
+From \emph{v1.0} of \cnjl, \texttt{code} accepts \texttt{:maths, :Julia, :C, :Cpp} as values for the \emph{lang} key argument.
+
+\begin{lstlisting}[language = Julia]
+function code(
+  c::Composition, lang=:maths;
+  name="composition"
+)
+  return get!(
+    c.code, lang, generate(c, name, Val(lang))
+  )
+end
+\end{lstlisting}
+
+\section{Explore, learn, and compose}
+\label{sec:xlc}
+
+It is far from practical to apply the internals of \cnjl step-by-step.
+We provide a user-friendly sequence of higher-level actions, \emph{explore}, \emph{learn}, and \emph{compose} that fits most of expected uses of this package.
+
+Each of those actions is semi-automated through default parameters and machine learning over a large collection of benchmarks \cite{baffier2022interpretable}.
+
+\subsection{Exploration}
+
+\begin{lstlisting}[language = Julia]
+function explore(
+  domains,
+  concept,
+  param=nothing;
+  search=:flexible,
+  complete_search_limit=10^6,
+  max_samplings=sum(domain_size, domains),
+  solutions_limit=floor(Int, sqrt(max_samplings)),
+)
+  if search == :flexible
+    search = sum(domain_size, domains) <
+      complete_search_limit ? :complete : :partial
+  end
+  return explore(Val(search), domains, concept,
+    param, solutions_limit, max_samplings
+  )
+end
+\end{lstlisting}
+
+\section{Future challenges}
+\label{sec:future}
+
 \begin{itemize}
-\item First unnumbered bulleted item which has no label and is indented
-from the left margin.
-\item Second unnumbered bulleted item.
-\item Third unnumbered bulleted item which has no label and is indented
-from the left margin.
+  \item Add new operations in existing layers
+  \item Add new layers
+  \item Use reinforcement learning
+  \item Cover more metrics
+  \item Apply ICN to other field than EFN for CBLS solvers
 \end{itemize}
-was produced by:
-\begin{verbatim}
-\begin{itemize}
-\item First item...
-\item Second item...
-\item Third item...
-\end{itemize}
-\end{verbatim}
 
-Numbered list is also provided in acmtog class file using the
-enumerate environment for example,
-\begin{enumerate}
-\item The attenuated and diluted stellar radiation.
-\item Scattered radiation, and
-\item Reradiation from other grains.
-\end{enumerate}
+\section{Problems, constraints, and compositions zoo}
+\label{sec:zoo}
 
-was produced by:
-\begin{verbatim}
-\begin{enumerate}
-\item The attenuated...
-\item Scattered radiation, and...
-\item Reradiation from other grains...
-\end{enumerate}
-\end{verbatim}
-\subsection{Illustrations (or figures)}
-\label{subsub:sec_Illus}
-The \verb|juliacon|  class file will cope with most of the positioning of
-your illustrations and you should not normally use the optional positional
-qualifiers on the \verb|figure|  environment that would override
-these decisions.
-\vskip 6pt
+The problems modelled, and the compositions extracted in this section are subject to future changes and improvements of the JuMP/JuCP packages.
+However, the keys ideas are presented here and examples will be updated accordingly within \jc in the future.
 
-%
-\begin{figure}[t]
-\centerline{\includegraphics[width=4cm]{juliagraphs.png}}
-\caption{This is example of the image in a column.}
-	\label{fig:sample_figure}
-\end{figure}
+\subsection{A few combinatorial models}
+\label{subsec:models}
 
-The figure \ref{fig:sample_figure} is taken from the JuliaGraphs
-organization \footnote{https://github.com/JuliaGraphs}.
+\begin{lstlisting}[language = Julia]
+function mincut(graph, source, sink, interdiction)
+  m = JuMP.Model(CBLS.Optimizer)
+  n = size(graph, 1)
+  separator = n + 1
 
-Figure captions should be \emph{below} the figure itself, therefore the
-\verb|\caption|  command should appear after the figure or space left for
-an illustration. For example, Figure 1 is produced using the following
-commands:
+  @variable(m, 0 <= X[1:separator] <= n, Int)
 
-\begin{verbatim}
-\begin{figure}
-\centerline{\includegraphics[width=20pc]{Graphics.eps}}
-\caption{An example of the testing process for a
-binary tree. The globa null hypothesis is tested
-first at level $\alpha$ (a), and the level of
-individual variables is reached last (d). Note
-that individual hypotheses can be tested at
-level $\alpha/4$ and not $\alpha/8$ as one might
-expect at first.}
-\label{sample-figure_2}
-\end{figure}
-\end{verbatim}
-Figures can be resized using first and second argument of
-\verb|\includegraphics|   command. First argument is used for modifying
-figure height and the second argument is used for modifying
-figure width respectively.
-\vskip 6pt
-Cross-referencing of figures, tables, and numbered, displayed
-equations using the \verb|\label|  and \verb|\ref|   commands is encouraged.
-For example, in referencing Figure 1 above, we used
-\verb|Figure~\ref{sample-figure}|
+  @constraint(m,
+    [X[source], X[separator], X[sink]] in Ordered()
+  )
+  @constraint(m, X in AllDifferent())
 
- \subsection{Tables}
-\label{subsub:sec_Tab}
-The \verb|juliacon|   class file will cope with most of the positioning of
-your tables and you should not normally use the optional positional qualifiers on the table environment which would override these
-decisions. Table captions should be at the top.
-\begin{verbatim}
-\begin{table}
-\tbl{Tuning Set and Testing Set}{
-\begin{tabular}{|l|l|c|c|}\hline
-Label & \multicolumn{1}{c|}{Description}
-& Number of Users &
-Number of Queries\\\hline
-Train70 & Training Data &
-\smash{\raise-7pt\hbox{70}} & 104\\
-\cline{1-2}\cline{4-4}
-Test70 & Testing Data I & & 105\\\hline
-Test30 & Testing Data II & 30 & 119\\\hline
-& Total & 100 & 328\\\hline
-\end{tabular}}
-\end{table}
-\end{verbatim}
+  obj(x...) = o_mincut(graph, x...; interdiction)
+  @objective(m, Min, ScalarFunction(obj))
 
-\begin{table}
-\tbl{Tuning Set and Testing Set}{
-\begin{tabular}{|l|l|c|c|}\hline
-Label & \multicolumn{1}{c|}{Description}
-& Number of Users &
-Number of Queries\\\hline
-Test 1 & Training Data &
-\smash{\raise-7pt\hbox{70}} & 104\\
-\cline{1-2}\cline{4-4}
-Test 2 & Testing Data I & & 105\\\hline
-Test 3 & Testing Data II & 30 & 119\\\hline
-& Total & 100 & 328\\\hline
-\end{tabular}}
-\end{table}
-\subsection{Landscaping Pages}
-\label{subsub:landscaping_pages}
-If a table is too wide to fit the standard measure, it may be turned,
-with its caption, to 90 degrees. Landscape tables cannot be produced
-directly using the \verb|juliacon|   class file because \TeX{} itself cannot
-turn the page, and not all device drivers provide such a facility.
-The following procedure can be used to produce such pages.
-\vskip 6pt
-Use the package \verb|rotating|   in your document and change the coding
-from
-\begin{verbatim}
-\begin{table}...\end{table}
-to
-\begin{sidewaystable}...\end{sidewaystable}
-and for figures
-\begin{figure}...\end{figure}
-to
-\begin{sidewaysfigure}...\end{sidewaysfigure}
-\end{verbatim}
+  return m, X
+end
+\end{lstlisting}
 
-environments in your document to turn your table on the appropriate
-page of your document. For instance, the following code prints
-a page with the running head, a message half way down and the
-table number towards the bottom.
-\begin{verbatim}
-\begin{sidewaystable}
-\tbl{Landscape table caption to go here.}{...}
-\label{landtab}
-\end{sidewaystable}
-\end{verbatim}
+\begin{lstlisting}[language = Julia]
+function golomb(n, L)
+  m = JuMP.Model(CBLS.Optimizer)
 
-\subsection{Double Column Figure and Tables}
-\label{subsub:double_fig_tab}
-For generating the output of figures and tables in double column
-we can use the following coding:
+  @variable(m, 0 <= X[1:n] <= L, Int)
 
-\begin{enumerate}
-\item For Figures:
-\begin{verbatim}
-\begin{figure*}...\end{figure*}
-\end{verbatim}
-\item For landscape figures:
-\begin{verbatim}
-\begin{sidewaysfigure*}...\end{sidewaysfigure*}
-\end{verbatim}
-\item For Tables:
-\begin{verbatim}
-\begin{table*}...\end{table*}
-\end{verbatim}
-\item For landscape tables:
-\begin{verbatim}
-\begin{sidewaystable*}...\end{sidewaystable*}
-\end{verbatim}
-\end{enumerate}
+  @constraint(m, X in AllDifferent())
+  @constraint(m, X in Ordered()) # optional
 
-\subsection{Typesetting Mathematics}
-\label{subsub:type_math}
-The \verb|juliacon| class file will set displayed mathematics with center to
-the column width, provided that you use the \LaTeXe{} standard of
-open and closed square brackets as delimiters.
-The equation
-\[
-\sum_{i=1}^p \lambda_i = (S)
-\]
+  # No two pairs have the same length
+  for i in 1:(n - 1), j in (i + 1):n
+    for k in i:(n - 1), l in (k + 1):n
+      (i, j) < (k, l) || continue
+      @constraint(m,
+        [X[i], X[j], X[k], X[l]] in DistDifferent()
+      )
+    end
+  end
 
-was typeset using the acmtog class file with the commands
+  # Add objective
+  @objective(m, Min, ScalarFunction(maximum))
 
-\begin{verbatim}
-\[
-\sum_{i=1}^p \lambda_i = (S)
-\]
-\end{verbatim}
+  return m, X
+end
+\end{lstlisting}
 
-For display equations, cross-referencing is encouraged. For example,
-\begin{verbatim}
-\begin{equation}
-(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
-\label{eq:samplevar}
-\end{equation}
-Equation~(\ref{eq:samplevar}) gives the formula for
-sample variance.
-\end{verbatim}
-The following output is generated with the above coding:
-\begin{equation}
-(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
-\label{eq:samplevar}
-\end{equation}
-Equation~(\ref{eq:samplevar}) gives the formula for
-sample variance.
+\subsection{Constraints and compositions}
+\label{subsec:constraints}
 
-\subsection{Enunciations}
-\label{subsub:enunciation}
-The \verb|juliacon|   class file generates the enunciations with the help of
-the following commands:
-\begin{verbatim}
-\begin{theorem}...\end{theorem}
-\begin{strategy}...\end{strategy}
-\begin{property}...\end{property}
-\begin{proposition}...\end{proposition}
-\begin{lemma}...\end{lemma}
-\begin{example}...\end{example}
-\begin{proof}...\end{proof}
-\begin{definition}...\end{definition}
-\begin{algorithm}...\end{algorithm}
-\begin{remark}...\end{remark}
-\end{verbatim}
-The above-mentioned coding can also include optional arguments
-such as
-\begin{verbatim}
-\begin{theorem}[...]. Example for theorem:
-\begin{theorem}[Generalized Poincare Conjecture]
-Four score and seven ... created equal.
-\end{theorem}
-\end{verbatim}
+Along with the magic-square model in Section \ref{sec:introduction}, the models in this zoo use the following constraints: \texttt{:all\_different}, \texttt{:dist\_different}, \texttt{:linear}, and \texttt{:ordered}.
 
-\begin{theorem}[Generalized Poincare Conjecture]
-Four score and seven years ago our fathers brought forth,
-upon this continent, a new nation, conceived in Liberty,
- and dedicated to the proposition that all men are
-created equal.
-\end{theorem}
+The \emph{AllDifferent} constraint ensures that all the values of a given configuration are unique.
 
-\subsection{Extract}
-\label{subsub:extract}
-Extract environment should be coded within
-\begin{verbatim}
-\begin{extract}..\end{extract}
-\end{verbatim}
+\begin{lstlisting}[language = Julia]
+function all_different(x;
+  X = zeros(length(x), 1), param=nothing, dom_size
+)
+  tr_in(Tuple([tr_count_eq_left]), X, x, param)
+  for i in 1:length(x)
+    X[i,1] = ar_sum(@view X[i,:])
+  end
+  return ag_count_positive(@view X[:,1]) |> (
+      y -> co_identity(
+        y; param, dom_size, nvars=length(x)
+      )
+    )
+end
+\end{lstlisting}
 
-\subsection{Balancing column at last page}
-\label{subsub:Balance}
-For balancing the both column length at last page use :
-\begin{verbatim}
-\vadjust{\vfill\pagebreak}
-\end{verbatim}
+\emph{DistDifferent} is constraint ensuring that \(|x[1] - x[2]| \neq |x[3] - x[4]|\) for any vector $x$ of size $4$.
 
-%\vadjust{\vfill\pagebreak}
+\begin{lstlisting}[language = Julia]
+function dist_different(x)
+  return abs(x[1] - x[2]) != abs(x[3] - x[4])
+end
+\end{lstlisting}
 
-at appropriate place in your \TeX{} file or in bibliography file.
+\emph{Linear} (also called \emph{Sum} or \emph{LinearSum} in the literature) Global ensures that the sum of the values of $x$ is equal to a given parameter $param$.
+Note that this version is a simplification of a linear sum of values with some coefficients.
 
-\section{Handling references}
-\label{subsub:references}
-References are most easily (and correctly) generated using the
-BIBTEX, which is easily invoked via
-\begin{verbatim}
-\bibliographystyle{juliacon}
-\bibliography{ref}
-\end{verbatim}
-When submitting the document source (.tex) file to external
-parties, the ref.bib file should be sent with it.
-\cite{bezanson2017julia}
+\begin{lstlisting}[language = Julia]
+function linear(x; param, dom_size)
+  return abs(sum(x) - param) / dom_size
+end
+\end{lstlisting}
+
+\emph{Ordered} is a constraint ensuring that all the values of $x$ are ordered (here in a decreasing order).
+
+\begin{lstlisting}[language = Julia]
+function ordered(x;
+  X = zeros(length(x), 1), param=nothing, dom_size
+)
+  tr_in(
+    Tuple([tr_contiguous_vals_minus]), X, x, param
+  )
+  for i in 1:length(x)
+    X[i,1] = ar_sum(@view X[i,:])
+  end
+  return ag_count_positive(@view X[:,1]) |> (
+      y -> co_identity(
+        y; param, dom_size, nvars=length(x)
+      )
+    )
+end
+\end{lstlisting}
+
+\section{Acknowledgements}
+\label{sec:acknowledgements}
+
+We're grateful to Ricardo \textsc{Rosa} and Pranshu \textsc{Malik} for the logo of \jc presented as Figure \ref{fig:juliaconstraints}.
 

### 2021-10-07 paper/header.tex

@@
 
-\title{CompositionalNetworks.jl: a scaling glass-box neural network for combinatorial functions learning}
+\title{CompositionalNetworks.jl: a scaling glass-box neural network to learn combinatorial functions}
 
@@
 \author[2]{Salvador \textsc{Abreu}}
-\affil[1]{IIJ Innovation Institute, Japan}
+\affil[1]{IIJ, Japan}
 \affil[2]{NOVA-LINCS, University of Évora, Portugal}
@@
 
-\keywords{Julia, Optimization, Constraint Programming, Neural Network, Scaling Learning}
+\keywords{Julia Language, Constraint Programming, Local Search, Metaheuristics, Neural Network, Metaprogramming, Scalable Machine Learning, Glass-Box Algorithm}
 
 \hypersetup{
-pdftitle = {CompositionalNetworks.jl: a scaling glass-box neural network for combinatorial functions learning},
+pdftitle = {CompositionalNetworks.jl: a scaling glass-box neural network to learn combinatorial functions},
 pdfsubject = {JuliaCon 2019 Proceedings},
 pdfauthor = {Jean-François \textsc{Baffier}, Khalil \textsc{Chrit}, Florian \textsc{Richoux}, Pedro \textsc{Patinho}, Salvador \textsc{Abreu}},
-pdfkeywords = {Julia, Optimization, Constraint Programming, Neural Network, Scaling Learning},
+pdfkeywords = {Julia Language, Constraint Programming, Local Search, Metaheuristics, Neural Network, Metaprogramming, Scalable Machine Learning, Glass-Box Algorithm},
 }

### 2021-10-07 paper/paper.tex

@@
 
-\hypersetup{colorlinks=true,linkcolor=blue,urlcolor=blue}
+\usepackage[section]{placeins}
+
+% \usepackage{float}
+
+\hypersetup{colorlinks=false,linkcolor=blue,urlcolor=blue}
 
@@
 \newcommand{\cblsjl}{\href{https://github.com/JuliaConstraints/CBLS.jl}{CBLS.jl}\xspace}
+\newcommand{\icnbjl}{\href{https://github.com/JuliaConstraints/ICNBenchmarks.jl}{ICNBenchmarks.jl}\xspace}
 
@@
   Interpretable Compositional Networks (ICNs) are a neural network variant for combinatorial function learning that allows the user to obtain interpretable results, unlike ordinary artificial neural networks.
-An ICN outputs a composition of functions that scale with the size of the input, allowing a learning phase on relatively small spaces.
+An ICN outputs a composition of functions that scales with the size of the input, allowing a learning phase on relatively small spaces.
   \cnjl is a pure Julia package that exploits the language's meta-programming, parallelism and multiple dispatch features to produce learned compositions in mathematical and programming languages such as Julia, C or C++.
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
   \centering
@@
   }
-
+  \hfill
   \subcaptionbox{EFSP landscape\label{sfig:efsp}}{
@@
     rate.}\label{fig:landscape}
-\end{figure}
+\end{figure*}
 
@@
 
-\subsection{A first C++ library}
-\label{subsec:cplusplus}
+% \subsection{A first C++ library}
+% \label{subsec:cplusplus}
 
@@
 
-As mentioned in Section \ref{sec:xlc}, \cnjl generates code for a direct use in Julia, but also exports usable code for both the GHOST(C++) and AdaptiveSearch (C)\footnote{This feature is a WIP at the moment this article is submitted, but is expected to be completed in the coming weeks.}.
+As mentioned in Section \ref{sec:use}, \cnjl generates code for a direct use in Julia, but also exports usable code for both the GHOST(C++) and AdaptiveSearch (C)\footnote{This feature is a WIP at the moment this article is submitted, but is expected to be completed in the coming weeks.}.
 The code exported in C++, that is the \emph{operations} of each neuron, is strongly inspired by our C++ library.
 
-\section{The example of the \jc framework}
-\label{sec:juliaconstraints}
-
-The \cnjl package is part of the \jc GitHub organization that proposes a first collection of Julia packages for Constraint-Based Local Search (CBLS), a subfield of Constraint Programming.
-
-The main goal of this framework is to provide a set of high level semi-automatic tools which strike a good compromise between efficiency and ease of modeling.
-
-\subsection{History of \jc}
-\label{subsec:history}
-
-The LocalSearchSolvers.jl package is a Constraint-Based Local Search framework started in Fall 2020 and inspired by other CBLS solvers such as GHOST (C++) and AdaptiveSearch (C) that allows users to tune their own solver.
-
-During the development of LocalSearchSolvers.jl, we decided to split the code and functionality of the original framework into several independent packages for ease of maintenance and in hopes of providing common tools for other constraint programming packages in Julia.
-
-During 2021, these tools were collected into the \jc ecosystem as follows
-\begin{itemize}
-  \item \cdjl: creation of discrete, continuous, and arbitrary domain (beta status)
-  \item \cjl: generation of constraints from a boolean concept or an error function. Also list usual constraints and their properties (beta status)
-  \item \cnjl: stable
-  \item \lssjl: A CBLS framework in Julia with built-in parallel and distributed scalability
-  \item \cblsjl: A MOI/JuMP wrapper for LocalSearchSolvers
-  \item \cmjl: list of CP models for \cblsjl and \lssjl
-\end{itemize}
+This work is part of a collection of articles that introduce the theoretical background of ICN \cite{richoux2020automatic}, a user-friendly implementation in Julia, and an extensive collection of benchmarks \cite{baffier2022interpretable}.
+Please note that the default parameters, and semi-automated learning parameters for advanced users, in \cnjl will improve along the results in \cite{baffier2022interpretable} computed from \icnbjl on an HPC cluster.
 
@@
 
+\FloatBarrier
+
+\section{The example of the \jc framework}
+\label{sec:juliaconstraints}
+
+The \cnjl package is part of the \jc GitHub organization that proposes a first collection of Julia packages for Constraint-Based Local Search (CBLS), a subfield of Constraint Programming.
+All packages fall under the MIT license.
+
+The main goal of this framework is to provide a set of high level semi-automatic tools which strike a good compromise between efficiency and ease of modeling.
+
+\subsection{History of \jc}
+\label{subsec:history}
+
+The \lssjl package is a Constraint-Based Local Search framework started in Fall 2020 and inspired by other CBLS solvers such as GHOST (C++) and AdaptiveSearch (C) that allows users to tune their own solver.
+
+During the development of LocalSearchSolvers.jl, we decided to split the code and functionality of the original framework into several independent packages for ease of maintenance and in hopes of providing common tools for other constraint programming packages in Julia.
+These tools were collected into the \jc ecosystem.
+
 The global state of the Constraint Programming ecosystem is presented in Figure \ref{fig:overview}.
-Beside the CBLS framework of \jc, other solvers are complete search methods, such as ConstraintSolver.jl\footnote{\url{https://github.com/Wikunia/ConstraintSolver.jl}}, CPLEXCP.jl\footnote{\url{https://github.com/dourouc05/CPLEXCP.jl}}, and SeaPearl.jl \cite{chalumeau2021seapearl}.
+Beside the CBLS framework of \jc detailed below, other solvers are complete search methods, such as ConstraintSolver.jl\footnote{\url{https://github.com/Wikunia/ConstraintSolver.jl}}, CPLEXCP.jl\footnote{\url{https://github.com/dourouc05/CPLEXCP.jl}}, and SeaPearl.jl \cite{chalumeau2021seapearl}.
+
+\paragraph*{State of the packages in \jc}
+
+\begin{itemize}
+  \item Stable packages
+  \begin{itemize}
+    \item \cnjl: internal tool to provide semi-automated error functions.
+Also, a standalone to learn compositions for combinatorial functions\footnote{At the time of submission, \cnjl is not yet release as \texttt{v1}.
+Some minor changes are expected from the results of the benchmarks in \cite{baffier2022interpretable} that will be submitted at the end of October.}
+  \end{itemize}
+  \item Beta packages (usable with frequent breaking changes)
+  \begin{itemize}
+    \item \cdjl: creation of discrete, continuous, and arbitrary domains
+    \item \cjl: generation of constraints from a boolean concept or an error function. Also, list usual constraints and their properties
+    \item \lssjl: A CBLS framework in Julia with built-in parallel and distributed scalability
+    \item \cblsjl: A MOI/JuMP wrapper for LocalSearchSolvers
+    \item \cmjl: list of CP models for \cblsjl and \lssjl
+  \end{itemize}
+
+\end{itemize}
+
+\subsection{The \emph{JuMP} ecosystem} % 3-6p
+\label{subsec:jump}
+
+JuMP is a modeling language accessible through its main package
+JuMP.jl \cite{DunningHuchetteLubin2017} and many supporting Julia packages. Beneath JuMP is an
+abstraction layer called MathOptInterface
+(MOI)~\cite{MathOptInterface-2021}. One can make a parallel
+between MOI/JuMP and FlatZinc/MiniZinc~\cite{nethercote2007minizinc},
+common modeling tools of the Constraint Programming community.
+Recently, the
+\href{https://github.com/dourouc05/ConstraintProgrammingExtensions.jl}{ConstraintProgrammingExtensions.jl}
+package that extends MOI/JuMP to the Constraint Programming world was
+made available. It is likely to become the recommended unified JuMP
+modeling extension for the Julia Constraint Programming solvers.
 
@@
 \label{subsec:example}
+
+\begin{figure}[t]
+  \centerline{\includegraphics[page=1, width=.75\columnwidth]{figs/magicsquare.pdf}}
+  \caption{A solved magic-square of size $3$. The \emph{magic sum} is equal to $15$. Image from Wikipedia's user \textsc{Phidauex}.}
+  \label{fig:magicsquare}
+\end{figure}
+
+Our CBLS framework uses three levels of modeling syntax: a \emph{raw} Julia syntax, and both MOI and JuMP syntaxes. In this article, we will only use the highest level one, the JuMP syntax.
 
@@
 
-\begin{figure}[t]
-  \centerline{\includegraphics[page=1]{figs/magicsquare.pdf}}
-  \caption{A solved magic-square of size $3$. The \emph{magic sum} is equal to $15$. Image from Wikipedia's user \textsc{Phidauex}.}
-  \label{fig:magicsquare}
-\end{figure}
-
-In \texttt{ConstraintModels.jl}, we use the following code to generate a magic-square instance.
+In \cmjl, we use the following code to generate a magic-square instance.
 
@@
 end
-
-# Create a magic square instance of size 4x4
-m, X = magic_square(4)
-# Optimize
-optimize!(m)
-# Output values
-@info values.(X)
 \end{lstlisting}
@@
 
-\cnjl extends the work of the first C++ library used in \cite{richoux2020automatic}.
-As mentioned in Section \ref{subsec:cplusplus}, that library used for prototyping was limited through several aspects: this, this, and that.
-Obviously, those limitations could have been lifted in the first library.
-However, we decided to use some features of the Julia language to ease provisioning flexibility and broader features to ICNs.
+As mentioned in Subsection \ref{subsec:icn}, \cnjl extends the work of the first C++ library used for prototyping in \cite{richoux2020automatic}.
+We decided to use some features of the Julia language to ease provisioning flexibility and broader features to ICNs.
 
-\paragraph*{Block about Julia} test
+\subsection{The Julia language}
+\label{subsec:julialang}
+
+The Julia language was introduced in~\cite{bezanson2017julia} and
+purports to be a new take on \emph{mathematical programming}.  Some
+key features of the language, according to the official website, which
+led us to consider it as an implementation vehicle:
+\begin{description}
+\item[Fast:] Julia was designed from the beginning for high
+  performance.  Julia programs compile to efficient native code for
+  multiple platforms via LLVM~\cite{lattner2004llvm}.
+\item[Dynamic:] It is dynamically typed, feels like a scripting
+  language, and has good support for interactive use.
+\item[Reproducible:] Reproducible environments make it possible to
+  recreate the same Julia environment every time, across platforms,
+  with pre-built binaries.
+\item[Composable:] It uses multiple dispatch as a paradigm, making it
+  easy to express many object-oriented and functional programming
+  patterns.
+\item[General-purpose:] It provides metaprogramming, asynchronous I/O,
+  debugging, logging, profiling, a package manager, and more.  One can
+  build entire Applications and Microservices in Julia.
+\item[Open source:] It is an open source project with over 1,000
+  contributors.  It is made available under the MIT license and the
+  source code is available on GitHub.
+\end{description}
 
@@
 A layer is defined by a non-empty collection of (possibly mutually exclusive) operations.
-The original ICN is composed of four layers: \emph{transformation} (exclusive), \emph{arithmetic}, \emph{aggregation}, and \emph{comparison}.
+Our basic ICN (for \efsp/\efop) is composed of four layers: \emph{transformation}, \emph{arithmetic} (exclusive), \emph{aggregation} (exclusive), and \emph{comparison} (exclusive).
+An illustration is given as Figure \ref{fig:model_icn}.
+The \texttt{Layer} structure stores these operations.
 
@@
 \end{lstlisting}
+
+The most complex layer is probably the \emph{transformation} layer.
+An operation (neuron) of this layer is defined over a vector of values and an optional parameter.
+As it is often simpler to define a transformation as a function over a specific index of the input vector, we provide meta-programming utility functions to generate vectorized methods: \texttt{lazy} and \texttt{lazy\_param}.
 
@@
 \end{lstlisting}
+
+When a concept is provided to an ICN, the user can provide an optional parameter input with the keyword argument \texttt{param}.
+A layer will be generated with or without the parametric operations (neurons) based on \texttt{param}'s value.
+Follow a shortened version of the transformation layer generator.
 
@@
 All methods required for the definition of additional layers are all available.
+An (unweighted) ICN is simply generated from a sequence of layers.
+After a learning phase using the a genetic algorithm from \texttt{Evolutionary.jl}\footnote{\url{https://github.com/wildart/Evolutionary.jl}}, this ICN can output a composition in various forms.
 
@@
 
-We store the composition of an ICN in three forms:
-\begin{itemize}
-  \item A collection of \texttt{Symbol}s per layer of the ICN that allows the generation of code as a string
-  \item The Julia object of type \texttt{Function} to apply the composition directly
-\end{itemize}
+Once the ICN is weighted, we store its  output as a composition with three forms.
+A collection of \texttt{Symbol}s per layer of the ICN that allows the generation of code as a pre-process in any language, assuming the operations encoding in that language are available.
+\cnjl currently supports Julia, C, and C++.
+
+A composition also stores a Julia object of type \texttt{Function} to apply the composition in a dynamic fashion.
 
@@
 
-The result output by an ICN is a composition, a mathematical function composed of several basic operations.
-The \texttt{code} function returns the definition of a composition in either a mathematical or programming language.
-From \emph{v1.0} of \cnjl, \texttt{code} accepts \texttt{:maths, :Julia, :C, :Cpp} as values for the \emph{lang} key argument.
+As mentioned above, an ICN outputs a composition, \ie{} a mathematical function composed of several basic operations.
+The \texttt{code} function returns the definition of a composition in either a mathematical or a programming language.
+From \texttt{v1} of \cnjl, \texttt{code} accepts \texttt{:maths} (default), \texttt{:Julia}, \texttt{:C}, \texttt{:Cpp} as values for the \texttt{lang} keyword argument.
 
@@
 
-\section{Explore, learn, and compose}
-\label{sec:xlc}
+\section{How to use \cnjl}
+\label{sec:use}
 
-It is far from practical to apply the internals of \cnjl step-by-step.
-We provide a user-friendly sequence of higher-level actions, \emph{explore}, \emph{learn}, and \emph{compose} that fits most of expected uses of this package.
+This section covers the essential steps to use ICN in a Julia environment. The actions required are few, which allow users to apply or export compositions effortlessly.
 
+\subsection{Installation and import}
+\label{subsec:install}
+
+\cnjl will keep compatibility with the latest stable and LTS releases of Julia\footnote{Note that at the time of writing, Julia has just released \texttt{v1.7}.
+Compatibility with last LTS will start from the next LTS release.}.
+We recommend the users to install the latest version of Julia\footnote{\url{https://julialang.org/downloads/}} and to call it with \texttt{julia -t auto} which launch Julia with all available threads on the local machine.
+Installing \cnjl from the Julia REPL is as simple as:
+
+\begin{lstlisting}[language = Julia]
+using Pkg
+Pkg.update()
+Pkg.add("CompositionalNetworks")
+\end{lstlisting}
+
+Alternatively, one can install from the package REPL (simply press `]' in a Julia session) in a single command.
+
+\begin{lstlisting}
+add CompositionalNetworks
+\end{lstlisting}
+
+Finally, loading the package is done by calling:
+\begin{lstlisting}[language = Julia]
+using CompositionalNetworks
+\end{lstlisting}
+
+Note that from Julia \texttt{v1.7}, calling \texttt{using} on uninstalled package will prompt the user to install it. Hence, the last command is the only one strictly required.
+
+\subsection{Explore, learn, and compose}
+\label{subsec:xlc}
+
+It is far from practical to apply the internals of \cnjl step-by-step. We provide a user-friendly sequence of higher-level actions that fits most of expected uses of this package:
+\begin{itemize}
+  \item \emph{explore} a search space according to a collection of variables domains, a metric, and a concept with an optional parameter
+  \item \emph{learn} the weights of an ICN on that space according to the same metric
+  \item \emph{compose} the elementary operations according to the weights of the ICN
+\end{itemize}
+
+The whole sequence can be executed through \texttt{explore\_learn\_compose}. That function accepts an increasingly large collection of (keyword) arguments:
+\begin{itemize}
+  \item \texttt{domains}: domains of the variables that define the training space
+  \item \texttt{concept}: the concept of the targeted constraint
+  \item \texttt{param}: an optional parameter of the constraint
+  \item \texttt{search}: either \texttt{:flexible},\texttt{:partial} or \texttt{:complete} search.
+Flexible search will use \texttt{search\_limit} and \texttt{solutions\_limit} to determine if the search space needs to be partially or completely explored
+  \item \texttt{global\_iter}: number of learning iterations
+  \item \texttt{local\_iter}: number of generations in the underlying genetic algorithm
+  \item \texttt{metric}: the metric to measure the distance between a configuration and known solutions
+  \item \texttt{pop\_size}: size of the population in the genetic algorithm
+  \item \texttt{complete\_search\_limit}: used in conjunction with \texttt{solution\_limit} to determine if a flexible search is set to complete or partial
+  \item \texttt{solutions\_limit}: see above
+  \item \texttt{sampler}: optional sampler function for a collection of configurations
+  \item \texttt{configurations}: optional collection of configurations to use as a training set
+  \item \texttt{memoize}: if true, use a memoized version of the metric
+\end{itemize}
 Each of those actions is semi-automated through default parameters and machine learning over a large collection of benchmarks \cite{baffier2022interpretable}.
 
-\subsection{Exploration}
-
 \begin{lstlisting}[language = Julia]
-function explore(
+function explore_learn_compose(
   domains,
@@
   param=nothing;
+  global_iter=Threads.nthreads(),
+  local_iter=100,
+  metric=:hamming,
+  pop_size=400,
   search=:flexible,
-  complete_search_limit=10^6,
-  max_samplings=sum(domain_size, domains),
-  solutions_limit=floor(Int, sqrt(max_samplings)),
+  complete_search_limit=1000,
+  solutions_limit=100,
+  sampler=nothing,
+  configurations=explore(
+    domains, concept, param;
+    search, complete_search_limit, solutions_limit
+  ),
+  memoize=true,
 )
-  if search == :flexible
-    search = sum(domain_size, domains) <
-      complete_search_limit ? :complete : :partial
-  end
-  return explore(Val(search), domains, concept,
-    param, solutions_limit, max_samplings
+  dom_size = maximum(domain_size, domains)
+  solutions, non_sltns = configurations
+  return learn_compose(
+    solutions,
+    non_sltns,
+    dom_size,
+    param;
+    local_iter,
+    global_iter,
+    metric,
+    pop_size,
+    sampler,
+    memoize,
   )
@@
 
+A long-term goal of \cnjl is to make the three first arguments optional too. We hope that such improvement will be used within \jc, for instance within the following more advanced examples.
+
+\subsection{Advanced examples}
+\label{subsec:advanced}
+
+Large scale use of \cnjl is made within \jc and appears in:
+\begin{itemize}
+  \item \cjl: the learning script can be found in \texttt{/src/learn.jl} and can be called as \texttt{learn\_from\_icn()}.
+  \item \icnbjl: an extensive collection of ICN benchmarks run on an HPC cluster. This package is not registered in the general registry of Julia packages.
+\end{itemize}
+
 \section{Future challenges}
@@
 
-\begin{itemize}
-  \item Add new operations in existing layers
-  \item Add new layers
-  \item Use reinforcement learning
-  \item Cover more metrics
-  \item Apply ICN to other field than EFN for CBLS solvers
-\end{itemize}
+We envisage several directions for future improvement of ICN within \cnjl. Follows a non-comprehensive list of the most challenging features.
+
+\paragraph*{New operations in existing layers} As the experiments have shown in \cite{richoux2020automatic,baffier2022interpretable}, some concepts are too complex to be covered with the current set of operations.
+The most direct solutions is to provide a broader collection of elementary operations.
+
+\paragraph*{New layers} Applied to \efsp/\efop or not, it is likely that adding some layers of operation could cover a wider spectrum of combinatorial functions (with some cost for the interpretability and computation time).
+
+\paragraph*{Reinforcement learning} Some limitations in the ICN parametrization and metrics could be solved at runtime with dynamic learning.
+In the \jc context, this could occur at the solver execution.
+
+\paragraph*{Apply ICN to other fields and problems} Contributions and suggestions are more than welcome!
 
@@
 However, the keys ideas are presented here and examples will be updated accordingly within \jc in the future.
+We welcome GitHub issues and pull requests.
 
@@
 \label{subsec:models}
+
+At the moment, \cmjl hosts more than a dozen models.
+Along with \emph{Magic Square} introduced in Section \ref{subsec:example}, we present models for two other classical optimization problems: \emph{Golomb ruler} and \emph{minimum cut}.
+The choice was made to cover different type of constraints, with error functions learned from \cnjl.
+
+Note that a model for the minimum cut problem in a network requires a matrix (graph) as input.
+Additionally, we add the optional argument of the interdiction of some links in the cut, which lead to an intractable problem called \emph{Network Interdiction} \cite{baffier2016parametric}.
 
@@
 
+In contrast, an instance of the Golomb ruler problem does not require specific dataset. However, this problem exists as both satisfaction and optimization variants.
+
 \begin{lstlisting}[language = Julia]
@@
   @constraint(m, X in AllDifferent())
-  @constraint(m, X in Ordered()) # optional
+
+  # optional for output readability
+  @constraint(m, X in Ordered())
 
@@
 
-Along with the magic-square model in Section \ref{sec:introduction}, the models in this zoo use the following constraints: \texttt{:all\_different}, \texttt{:dist\_different}, \texttt{:linear}, and \texttt{:ordered}.
+Along with the magic-square model in Section \ref{sec:introduction}, the models in this zoo use the following constraints: \texttt{:all\_different}, \texttt{:dist\_different}, \texttt{:linear}, and \texttt{:ordered}\footnote{The code provided here will be updated following the results of the benchmark in \cite{baffier2022interpretable}.
+In particular, for \texttt{:dist\_different} and \texttt{:linear}, the current code is a hand-written function.}.\newline
 

### 2021-10-07 dev/search_index.js
build based on 1c7ea1a

@@ -1,3 +1,3 @@
 var documenterSearchIndex = {"docs":
-[{"location":"domain/#ConstraintDomains.jl-(dependency)","page":"ConstraintDomains.jl","title":"ConstraintDomains.jl (dependency)","text":"","category":"section"},{"location":"domain/","page":"ConstraintDomains.jl","title":"ConstraintDomains.jl","text":"Currently only discrete domains are supported using the following function. ","category":"page"},{"location":"domain/","page":"ConstraintDomains.jl","title":"ConstraintDomains.jl","text":"ConstraintDomains.domain","category":"page"},{"location":"domain/#ConstraintDomains.domain","page":"ConstraintDomains.jl","title":"ConstraintDomains.domain","text":"domain()\n\nConstruct an EmptyDomain.\n\n\n\n\n\ndomain(a::Tuple{T, Bool}, b::Tuple{T, Bool}) where {T <: Real}\ndomain(intervals::Vector{Tuple{Tuple{T, Bool},Tuple{T, Bool}}}) where {T <: Real}\n\nConstruct a domain of continuous interval(s). ```julia d1 = domain((0., true), (1., false)) # d1 = [0, 1) d2 = domain([ # d2 = 0, 1) ∪ (3.5, 42, (1., false),     (3.5, false), (42., true), ])\n\n\n\n\n\ndomain(values)\ndomain(range::R) where {T <: Real, R <: AbstractRange{T}}\n\nConstruct either a SetDomain or a RangeDomain.\n\nd1 = domain(1:5)\nd2 = domain([53.69, 89.2, 0.12])\nd3 = domain([2//3, 89//123])\nd4 = domain(4.3)\n\n\n\n\n\n","category":"function"},{"location":"comparison/#Comparison-Layer","page":"Comparison","title":"Comparison Layer","text":"","category":"section"},{"location":"comparison/","page":"Comparison","title":"Comparison","text":"The comparison layer of our basic ICN can be constructed using comparison_layer(param=false). All operations are mutually exclusive.","category":"page"},{"location":"comparison/","page":"Comparison","title":"Comparison","text":"CompositionalNetworks.comparison_layer","category":"page"},{"location":"comparison/#CompositionalNetworks.comparison_layer","page":"Comparison","title":"CompositionalNetworks.comparison_layer","text":"comparison_layer(param = false)\n\nGenerate the layer of transformations functions of the ICN. Iff param value is set, also includes all the parametric comparison with that value. The operations are mutually exclusive,
[…]

### 2021-10-08 paper/paper.tex
Changes to solve affiliation issues
Fix a missing file (for build)

@@
 These constraints are mostly formulated as concepts, boolean functions indicating whether each constraint is respected or not.
-A solution (sometimes called a satisfactory solution, depending on the domain) is a configuration that respects all the constraints.
+A solution (sometimes called a satisfying solution, depending on the field) is a configuration that respects all the constraints.
 
@@
 
-This work, consists in a library that implements ICNs in the Julia programming language \cite{bezanson2017julia}.
+This work consists in a library that implements ICNs in the Julia programming language \cite{bezanson2017julia}.
 Although we present a direct application of ICNs in the following subsection, this neural network framework is simple to extend to learn other combinatorial and non-combinatorial functions.
@@
 An (unweighted) ICN is simply generated from a sequence of layers.
-After a learning phase using the a genetic algorithm from \texttt{Evolutionary.jl}\footnote{\url{https://github.com/wildart/Evolutionary.jl}}, this ICN can output a composition in various forms.
+After a learning phase using a genetic algorithm from \texttt{Evolutionary.jl}\footnote{\url{https://github.com/wildart/Evolutionary.jl}}, this ICN can output a composition in various forms.
 
@@
 
-Note that from Julia \texttt{v1.7}, calling \texttt{using} on uninstalled package will prompt the user to install it. Hence, the last command is the only one strictly required.
+Note that from Julia \texttt{v1.7}, calling \texttt{using} on an uninstalled package will prompt the user to install it. Hence, the last command is the only one strictly required.
 
@@
 
-The \emph{AllDifferent} constraint ensures that all the values of a given configuration are unique.
+The \emph{AllDifferent} constraint ensures that all the values of a given configuration are unique. This constraint is representative of Constraint Programming and is used in a large amount of models.
 
